### 2020-11-13 NSF2020/Description/broaderimpacts.tex
broader impacts edited

\section{Broader Impacts of the Proposed Work}
\label{sec:BroaderImpacts}

The proposed mathematical analysis, modeling, and numerical algorithms
will transform our understanding of the collective dynamics of
amphiphilic particles such as (1) their self-assembly into micelles and
bilayers, (2) the material properties of such self-assembly, and (3) the
interaction between these building blocks. In addition to biophysical
applications, amphiphilic Janus particles have recent popularity in the
fabrication of smart materials. The proposed work will have a
transformative impact on precision design for specific mechanical
properties of materials made of amphiphilic nanoparticles. An important
component of this proposal is the interdisciplinary education and
training of both undergraduate students, graduate students, and
postdoctoral fellows. The combination of mathematical modeling,
analysis, and scientific computing in this project provides a compelling
example of the importance of mathematics in biophysics and engineering
applications. The concepts and methods described here go beyond the
context of amphiphilic particles and lipid molecules. They extend to
other problems featuring microscopic phase separation that leads to
formation of mesoscopic domains. This situation arises, for example, in
biological development in systems biology. The methods developed here
have the potential to impact those and other related areas in
biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The proposed research will have an immediate impact on undergraduate and
graduate education. At the undergraduate level, the project will include
and support undergraduate researchers from Fordham University. The
supported summer researchers will gain valuable first-hand
interdisciplinary experience in mathematical modeling and computation.
We will incorporate research into teaching, especially in differential
equation and programming courses. PI YNY will work on simple modeling of
self-assembly problems with both undergraduate and high school students
from Newark Science Park High School. PI YNY has a track record of
working with local high school students for their research projects
before they apply for colleges. PI BQ will work with undergraduate and
high school students through the Undergraduate Research Opportunity
Program and Young Scholar's Program as he has done in the past.

Over the past eight years, PI RR has included many undergraduates in the
execution of research initiatives and coauthoring publications. He has a
track record for supporting underrepresented groups, being called on by
the Collegiate Science and Technology Entry Program (CSTEP) to make
opportunities for promising students, for example. He has mentored two
Clare Boothe Luce Scholars, one a U.S.~Marine Corp veteran, and a high
school student in the NYU GSTEM program.

The PIs will capitalize on the synergy between both FSU's doctoral
program in Scientific Computing, NJIT's doctoral program in Mathematical
Sciences, and Fordham's undergraduate focus in the Department of
Mathematics. Promising, young undergraduate scientists from the Bronx
community will find a natural pipeline into graduate studies by working
directly or indirectly on this project. At the graduate-level, PI YNY
expects to train and support a PhD student for two more years. PI BQ
will advise a postdoctoral fellow for two years. The PIs will foster
vertical integration between the senior personnel, their collaborators,
postdocs, doctoral, and Bachelor students, enhancing the learning
environment.

%The PIs expect
%to involve undergraduate students in this project through
%their continuing and active involvement in undergraduate advising and research mentoring:
%YNY has mentored undergraduate students as a co-Investigator in CSUMS: Research and Education
%in Computational Mathematics for undergraduates in the Mathematical 
%Sciences at NJIT (funded by NSF) and the lecturer for Capstone Applied
%Mathematics Lab at NJIT (also funded by NSF).
%YNY has track records in involving undergraduate students in research
%that emphasizes both numerical computations and desktop experiments.
%
%such as MATH 340: Advanced Numerical Methods.
%Simple MATLAB codes have been used as teaching tools to show students how to use MATLAB
%to simulate the slender-body equations for an elastic filament in Stokes flow.
%
%In the past year YNY has been mentoring Ufuomaefe Ogbe, a high school student
%from Newark. Ufuomaefe is an African American who 
%worked with PI on a simplified model for vesicles in flow. Based
%on what he learned about modeling and programming, 
%he has applied to the Applied Mathematics/Computer Science programs at both NJIT
%and Rutgers/Newark.

\section{Intellectual Merit}
A central theme of the project is the mathematical development and
analysis of the physical model for interaction between many amphiphilic
particles. The model formulates the interaction potential through a
screened Laplace equation boundary value problem possessing the physical
properties like non-additivity and decay of realistic hydrophobic
attraction. Colloidal systems collectively self-assemble into bilayer
morphologies, and we analyze the elastic properties of these amphiphilic
particle ensembles. This allows us to interpret the Helfrich free energy
in terms of hydrophobic interactions and specific molecular
characteristics. We extend the capabilities of the boundary integral
equation and time stepping methods to stably perform large particle
number simulations in three dimensions. Results from these computations
will be used to compare collective amphiphilic against experiments and
to study the optimal design of three-dimensional functional materials.