### 2020-11-15 NSF2020/Description/broaderimpacts.tex
RJR educational impact

\section{Broader Impacts}
\label{sec:BroaderImpacts}

The proposed mathematical analysis, modeling, and numerical algorithms
will transform our understanding of the collective dynamics of
amphiphilic particles such as (1) their self-assembly into micelles and
bilayers, (2) the material properties of such self-assembly, and (3) the
interaction between these building blocks. In addition to biophysical
applications, amphiphilic Janus particles have recent popularity in the
fabrication of smart materials. The proposed work will have a
transformative impact on precision design for specific mechanical
properties of materials made of amphiphilic nanoparticles. An important
component of this proposal is the interdisciplinary education and
training of both undergraduate students, graduate students, and
postdoctoral fellows. The combination of mathematical modeling,
analysis, and scientific computing in this project provides a compelling
example of the importance of mathematics in biophysics and engineering
applications. The concepts and methods described here go beyond the
context of amphiphilic particles and lipid molecules. They extend to
other problems featuring microscopic phase separation that leads to
formation of mesoscopic domains. This situation arises, for example, in
biological development in systems biology. The methods developed here
have the potential to impact those and other related areas in
biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The topics of this proposal have application in robotics, machine learning,
and engineering. To foster training in mathematical sciences, the project
will include undergraduate researchers (URs). The proposal supports one UR
per year from the lead institution, and will include one or two additional
UR from actively supported in-house programs

The URs will work as a team for eight weeks in the summer. They will
collaborate directly with the senior personnel, have tutorials in topics
such as numerical quadrature and integral equations, and train in
mathematical writing and presentation. To maximize vertical integration,
and to the extent possible, the PIs and personnel will travel to 
New York for one or two weeks in the summers. Fordham University 
recently constructed a seminar/collaboration space specifically
for collaborative research, and URs will base there activitied from here.

As a requirement, the URs will write quality summary reports and give
presentations at the their respective research symposia. The work will be
considered a success if the URs also participate in a national conference.
The best possible outcome is if URs coauthor a publication.

There are two ways the proposal will meaningfully engage the community.
To be more impactful, students from the Bronx High School of Science
(PI RR) and Newark Science Park High School (PI YNY) will be encouraged
to join the research team. PI BQ will work with undergraduate and high
school students through the Undergraduate Research Opportunity Program
and Young Scholar's Program, as he has done in the past. Secondly, the
PIs will prioritize students who would most benefit from the grant activity.
We specifically target students whose socio-economic background prevents
them from participating in out-of-state research experiences.

Innovations in teaching: we will create several modules to include
in our courses. These modules will introduce topics from numerical linear
algebra and optimization, and be based on the ideas of the proposal.
Additionally, we and our UR collaborators will give remote, guest lectures
in each others' courses and undergraduate seminars.

%The proposed research will have an immediate impact on undergraduate and
%graduate education. At the undergraduate level, the project will include
%and support undergraduate researchers from Fordham University. The
%supported summer researchers will gain valuable first-hand
%interdisciplinary experience in mathematical modeling and computation.
%We will incorporate research into teaching, especially in differential
%equation and programming courses. PI YNY will work on simple modeling of
%self-assembly problems with both undergraduate and high school students
%from Newark Science Park High School. PI YNY has a track record of
%working with local high school students for their research projects
%before they apply for colleges. PI BQ will work with undergraduate and
%high school students through the Undergraduate Research Opportunity
%Program and Young Scholar's Program as he has done in the past.

%Over the past eight years, PI RR has included many undergraduates in the
%execution of research initiatives and coauthoring publications. He has a
%track record for supporting underrepresented groups, being called on by
%the Collegiate Science and Technology Entry Program (CSTEP) to make
%opportunities for promising students, for example. He has mentored two
%Clare Boothe Luce Scholars, one a U.S.~Marine Corp veteran, and a high
%school student in the NYU GSTEM program.

%The PIs will capitalize on the synergy between both FSU's doctoral
%program in Scientific Computing, NJIT's doctoral program in Mathematical
%Sciences, and Fordham's undergraduate focus in the Department of
%Mathematics. Promising, young undergraduate scientists from the Bronx
%community will find a natural pipeline into graduate studies by working
%directly or indirectly on this project. At the graduate-level, PI YNY
%expects to train and support a PhD student for two more years. PI BQ
%will advise a postdoctoral fellow for two years. The PIs will foster
%vertical integration between the senior personnel, their collaborators,
%postdocs, doctoral, and Bachelor students, enhancing the learning
%environment.

%The PIs expect
%to involve undergraduate students in this project through
%their continuing and active involvement in undergraduate advising and research mentoring:
%YNY has mentored undergraduate students as a co-Investigator in CSUMS: Research and Education
%in Computational Mathematics for undergraduates in the Mathematical 
%Sciences at NJIT (funded by NSF) and the lecturer for Capstone Applied
%Mathematics Lab at NJIT (also funded by NSF).
%YNY has track records in involving undergraduate students in research
%that emphasizes both numerical computations and desktop experiments.
%
%such as MATH 340: Advanced Numerical Methods.
%Simple MATLAB codes have been used as teaching tools to show students how to use MATLAB
%to simulate the slender-body equations for an elastic filament in Stokes flow.
%
%In the past year YNY has been mentoring Ufuomaefe Ogbe, a high school student
%from Newark. Ufuomaefe is an African American who 
%worked with PI on a simplified model for vesicles in flow. Based
%on what he learned about modeling and programming, 
%he has applied to the Applied Mathematics/Computer Science programs at both NJIT
%and Rutgers/Newark.

\section{Intellectual Merit}
A central theme of the project is the mathematical development and
analysis of the physical model for interaction between many amphiphilic
particles. The model formulates the interaction potential through a
screened Laplace equation boundary value problem possessing the physical
properties like non-additivity and decay of realistic hydrophobic
attraction. Colloidal systems collectively self-assemble into bilayer
morphologies, and we analyze the elastic properties of these amphiphilic
particle ensembles. This allows us to interpret the Helfrich free energy
in terms of hydrophobic interactions and specific molecular
characteristics. We extend the capabilities of the boundary integral
equation and time stepping methods to stably perform large particle
number simulations in three dimensions. Results from these computations
will be used to compare collective amphiphilic against experiments and
to study the optimal design of three-dimensional functional materials.