### 2021-11-10 NSF2021/Description/specificaim1.tex
RJR - SP1, Prelim, new figure folders

\section{Proposed Research}
\label{sec:proposed-work}
%%
\begin{center}
  \begin{figure}
      \includegraphics[width=1\textwidth]{figures/SpecificAim1/Domain.pdf}
  \end{figure}
\end{center}

\subsection{Specific Aim 1: Measuring material properties of amphiphile
  self-assembly}
\label{subsec:specific_aim_1}

The proposed research aims to provide fundamental insight into the
self-assembly dynamics of amphiphilic particles. These results will
facilitate optimal design of smart materials by tuning the geometry and
properties of the amphiphilic particles.
Specific Aim 1 achieves this through a novel mathematical model
for particle self-organization driven by long-range interfacial forces.

\subsubsection{Problem formulation}
The mathematical formulation is a nonlinear system 
for the dynamics of a collection of particles. The 
interactions come from a system of linear partial differential equations
(PDE) and comprise hydrodynamic interactions and hydrophobic interactions.
The formulation prevents particle collisions through excluded volume
potentials.

The hydrodynamic interactions come from the mobility problem for rigid
particles immersed in a viscous solvent. Assuming inertial terms are
negligible, the particle motion is goverened by the Stokes equations
\begin{alignat}{3}
\label{eq:stokes1}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
\label{eq:stokes2}
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
\label{eq:stokes3}
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity and $p$ is the pressure of the solvent,
$\uu_\infty$ is the background flow, and $\mu$ is the constant
solvent viscosity. The domain
$\Omega$ is the fluid region surrounding the particles.
Since the particles are rigid, the solvent velocity 
satisfies a no-slip boundary condition for a rigid body motion 
%\begin{align}
%\label{eq:rigid_bc}
%  \uu(\xx) = \vv_i + \omega_i  (\xx - \aa_i)^\perp, \quad
%    \xx \in \Gamma_i,\qquad  i=1,\ldots,N_p,
%\end{align}
on the particle boundary $\Gamma_i$
with translational velocity $\vv_i$ and
angular velocity $\omega_i$.
Given imposed forces acting on each particle,
the \emph{mobility problem} consists of finding
translational and angular velocities so that 
viscous fluid forces balance the imposed forces.

The hydrophobic interactions come from the tendency 
of particles to minimize the free energy of the structure of
the surrounding water molecules.  
Experimental measurements suggest that the free energy functional
take the form 
\begin{equation}
\label{eq:free_energy}
F[u] = C \int_{\Omega}  \rho |\nabla u|^2 + \rho^{-1} f(u) \,dx
\end{equation}
where
$u$ is an order parameter for the structure of water,
$\rho$ is a decay length,
$C$ is a constant, and $f(u)$ is a potential.
Hydrogen-bond persistence times are on the order of picoseconds \cite{MaGa13}.
Since waters reorient 
on a time scale much smaller than for the particle motion, 
then at each moment in time $u$ minimizes $F[u]$ and satisfies 
\begin{equation}
\label{eq:SL}
-\rho^2 \Delta u + f'(u) = 0  \text{ in } \Omega,\quad u = g
\text{ on } \Omega.
\end{equation}
The boundary condition $g$ that imposes 
the water structure at the water-hydrocarbon interfaces.
It is a material label that is transported with the particle motions.

The coupling between the Stokes equations \eqref{eq:stokes1}-\eqref{eq:stokes3}
and the semilinear elliptic equation \eqref{eq:SL}
comes from the first variation of the energy functional.
The particles minimize the free energy $F[u]$
by altering the location of hydrophobic interfaces. 
It is possible to calculate the rate of change of $F[u]$
using variation of the domain~\cite{Bandle2015, Schiffer1954, Grinfeld2010}.
Carrying out this variation yields a 
%The hydrophobic forces are torques are
%\begin{equation}
%  \label{eq:hydrophobicAttraction}
%  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu \, \dif s, 
%    \quad 
%    T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot
%    ({\bf T} \cdot \nnu) \dif s.
%\end{equation}
surface force density 
\begin{align}
  \label{eq:stress}
\mathbf{T}
= C \left[ \rho^{-1} f(u) \mathbf{I}
  + \rho  \left(|\nabla u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)\right].
\end{align}
The imposed forces and torques from the integration of $\mathbf{T}$ along the
particle boundary.
The density \eqref{eq:stress}  was first introduced
by~\cite{Fu2018_SIAM} and is the higher-dimensional analogue of
disjoining pressure calculated by
\cite{MaRa76, ErLjCl89, KoNa15, Nagle17, KUZMIN2005}.
It is the mathematical ingredient responsible for forming particle
aggregates that sequester their hydrophobic surface regions.

By solving the above system, we obtain translational and angular
velocities of the many-body system. 
A second-order Adams-Bashforth scheme updates the particle positions and orientations.

The formulation presents a number of
mathematical and numerical challenges. For one, the domain is constantly
changing so that a new boundary value problem must be solved at each time step.
The particle boundaries are nearly touching and this requires
high-order numerical schemes to accurately calculated forces.
The domain is unbounded and there are issues with accurately
calculating the velocity in the far-field. 
Finally, the numerical routine must handle
large system sizes efficiently when their are many particles
is large.

%and
%\begin{equation}
%  \label{eq:force}
%  \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i,\quad 
%  \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot (\ssigma \cdot \nnu) \,
%  \dif s = T_i, \qquad i=1,\ldots,N_b,
%\end{equation}
%where
%$\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T \right)$
%is the hydrodynamic stress tensor (pressure tensor) and $\nnu$ is the
%particle outward normal.

\subsubsection{Janus-particle self-assembly} 
\begin{wrapfigure}[14]{l}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{figures/SpecificAim1/multilamelar.pdf}
\caption{\label{fig:multilamelar}
  Experimentalists can transform unilamellar vesicles to multilamellar
  vesicles by altering the quality of the solvent
    \cite{deMeetal21}.
  In \eqref{eq:SL}, the solvent quality enters through the potential
  $f(u)$ and boundary conditions. 
}
\end{wrapfigure}
A number of morphologies arise from the theory. In the linear response case,
take $f(u) = u^2$ so that water structure has its lowest potential value
at $u = 0$, representing bulk water. The solutions of \eqref{eq:SL} have
a boundary layer structure and decay to 0 in the bulk with the decay
length $\rho$.

Consider a polar boundary condition where
$g_i(\mathbf{x}) = (\pi r)^{-1/2} \cos (\theta_i(\mathbf{x}))$
where $\theta_i$ is the angle formed by $\mathbf{x}$, the particle
center $\mathbf{a}_i$, and the particle director $\mathbf{d}_i$.
The normalization $(\pi r)^{-1/2}$ here and others below
provide for a uniform surface energy $\int_{\Gamma_i} g_i^2 \,ds = 1$
when the particles are disks of radius $r$.
The boundary condition is odd along the particle axis 
making the particle head repel the tail.
We simulate the dynamics of 100 particles with random initial position
and orientation. The particles rapidly form chains with their
directors perpendicular to the length of the chain. The equilibrium
structure consists particles positioned on a square grid with alternating
orientations.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A5.png}\\
    Figure 2. Self assembly for a polar boundary condition
    \label{fig:self-assemblyA}
\end{center}

Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and are accounted for as follows.
The hydrophilic side takes the value $u = 0$ since the
apolar headgroup of a lipid, for example, does not alter the structure
of adjacent waters. The hydrophobic side takes the value $u > 0$. 
The interaction between particles is attractive. Using
$(3\pi )^{-1/2}(1 + \cos(\theta_i(\mathbf{x})))$
as a simple modification to the previous boundary condition 
produces two-dimensional micelles and bilayers \cite{Fu2018_SIAM}.
Researchers have asked us if the phenomenon is robust. To show
that it is, we use 
\begin{equation}
\label{eq:vonMises}
g_i(\mathbf{x}) =
\sqrt{
  \frac{\exp( \kappa_i(\theta_i(\mathbf{x}) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}
\end{equation}
instead where the radicand is the von Mises 
distribution with director angle $\theta^0_i$ and Bessel function
$I_0$ of order zero. The square of this function behaves as a normal
distribution with variance like $1/\kappa_i$. To replicate
experimental conditions, the concentration $\kappa_i$ and radii
$r_i$ are randomized. Using the same initial configuration, the
particles form small groups. The small groups combine to form bilayers.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B4.png}\\
    Figure 3. Self assembly for a amphiphilic boundary condition
    \label{fig:self-assemblyA}
\end{center}

Multilamelar bilayers arise with an increase by shifting the boundary condition
further. Let $g_i(\mathbf{x}) = (9\pi )^{-1/2}(2 + \cos(\theta_i(\mathbf{x})))$.
This gives a hydrophobic particle with greater intensity of hydrophobicity
at $\theta_i = 0$ than at $\theta_i = \pi$.  The initial self-assembly
is similar to that of the amphiphilic particle case, except that the bilayers
do not stay well separated.  Rather, the bilayers stack on top of each
other as a consequence of the interfacial tension of exposed particle heads. 
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C4.png}\\
    Figure 3. Self assembly for hydrophobic particles with bias.
    \label{fig:self-assemblyA}
\end{center}

In summary, the hydrophibic interaction-mobility problem model yields
rich morphologies that are valuable in the synthesizing multiphasic and
anisotropic colloids. The goal is to obtain numerically aided 
syntheses and assembly of soft matter materials.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-change to amplify colloid
utility for a wide range of  spanning environmental remediation to drug
delivery. The main experimental challenges include the ability to
decouple particle surface properties.
The main modeling challenge is to translate the phenomenological
parameters, e.g. boundary conditions, into measurable quantities
such as contact angles. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions. In Specific Aim 3, we
discuss how to include ions and surface charge.

\subsubsection{Hydrophobic forces as capillary wetting}
A natural generalization of the linear response model is 
to use a fourth-order, tilted double-well potential
$f(u) = a(u-u_0)^2(u-u_1)^2 + bu$.
Motivated by capillary wetting, soft matter physicist Gerhard Gompper 
introduced a double-well model for one-dimensional hydrophobic attraction
\cite{GoHaKo94}. The two-dimensional attraction we propose
is more complicated and includes extra line tension effects
not present in the one-dimensional theories.

Here, the order parameter $u$ is the mean number of nearest
neighbors per water molecule.
The value $u_0 = 4$ gives a local minimum for the ideal
tetrahadral water network found at hydrophobic surfaces.  
The value $u_1 \sim 4.4$ gives the absolute minimum
in the bulk where there is a mixture of four and five-coordinated
waters. The equation \eqref{eq:SL} takes the form of a
steady-state Allen-Cahn equation. 
When particles are well-separated, the energetic cost for filling
the interstitial region with $u_0$-phase water is prohibitive
and there is a phase shift from the $u_0$ to $u_1$ water order
---the attraction landscape is more
or less identical to the linear response theory. Conversely, when
particles are proximal, it becomes energetically favorable
to coalesce $u_0$-phases. This introduces a line tension
for the phase transition.

One potential application of the Allen-Cahn-type formulation is to
explaning measured hydrophobic attraction profiles.
The experimental data on hydrophobic attraction between 
is well-established and the literature contains data for a number of
solvents and hydrocarbon species. The experimental measurements
are well-fit by a double exponential curve in the 10 to 100 nanometer range
of separation.  Below 10 nm, however, there is an amplification of force
and the attraction follows a much steeper force law not accounted
for by past phenomenoloical theories \cite{Lin2005}.

It may be the case that the double-well formulation recreates the
long and short-range interactions.  The advantage
is that the two-dimensional simulations can replicate the
experimental setup better than prior, one-dimensional theoretical work 
Some of the model parameters directly connect 
to long-range decay lengths and energy cost of creating hydrophobic-water
interfaces with established ranges of values. Other parameters, like the
well structure, are fit to the measured data. These base-line simulations
will allow us to move on to provide more accurate colloidal suspension
simulations.

Other approaches will be required if the modified phenomenological theory 
does not recreate the measured force curves.
We can rule out convection in the water structure since  
the time scale of particle motion is much larger than
the hydrogen-bond lifetimes. One approach is to consider Gaussian fluctuation.
Here, we expand the free energy functional around the saddle point solution.
The fluctuation energy then involves calculation of the eigenfunctions of
the second variation of the free energy. This accounts for the fact
that experimental measurements are carried out at finite 
temperatures, where thermal fluctuations should ideally also be considered.

%
% The following commented out Nov 6, 2021.
%
%
\begin{comment}
The goal of Specific Aim 1 is to characterize the material properties of
many-body, self-assembled amphiphiles.  For amphiphiles assembled into
bilayers, these properties are described by membrane continuum
mechanics.  Our goal is to map the parameters of the particle-based
model onto the elastic moduli from continuum theory.  Results from this
goal will facilitate simulators to use the hydrophobic attraction force
calculations to model bilayers with specific composition. These
calculations have provably less computational complexity than those of
molecular dynamics simulations and possess the molecular granularity
lacking from continuum models.

Hamm and Kozlov (HK) pioneered the modern theory of membrane continuum
mechanics~\cite{Hamm2000}, and their theory is widely used to describe
biological phenomena, including fission \cite{FrEsAkSh15, Maetal15,
PhysRevE.79.031926}, fusion \cite{ChKo08,
KoKo2002,Kuzmin7235,Aeffner2012}, poration~\cite{Gaetal20}, phase
boundaries, and interaction with inclusions~\cite{SeLeMaEg17,Saetal20,
Pietal20}. These phenomena require resolution of the internal structure
of the membrane.  Recently, there has been a revival of interest in the
HK theory as the quadratic assumption for the elasticity energy density
has caused researchers to question the applicability of the theory for
large curvatures~\cite{PhysRevLett.117.188102, ARGUDO20161619}.
%
\begin{wrapfigure}[11]{l}{0.47\textwidth}
\centerline{\includegraphics[width=0.46\textwidth]{Figures/Deformations.pdf}}
  \vspace{-5pt}
\caption{\label{fig:deformations} \footnotesize Sketch of the HK
 membrane model \cite{Hamm2000}.}
\end{wrapfigure}
%
This proposal will develop much-needed mathematical analysis to resolve
these controversies due to the assumptions in the HK theory.

\sloppy
The HK framework assumes a three-dimensional lipid monolayer where the
internal structure consists of straight fibers that represent elongated
hydrocarbon chains.  The elastic energy density ${\cal W}$ is quadratic
in the Green-Lagrange strain tensor for this striated, internal
structure. This energy density decomposes into four, fundamental, and
independent deformations (Figure~\ref{fig:deformations}): splay ($\Div
{\bf n}$), twist ($\Curl {\bf n}$), saddle splay ($\det \nabla {\bf
n}$), and tilt ${\bf t}={\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where
${\bf N}$ is the unit surface normal;
\begin{equation}
\label{ansatz3}
{\cal W} \equiv \int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div {\bf n} + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl {\bf n})^2 + \KG  \det \nabla {\bf n} + \tfrac{1}{2}\KTH |{\bf t}|^2 \,dA.
\end{equation}
Here, the deformations come with elastic coefficients: the bending
modulus $\KB$, twist modulus $\KT$, saddle-splay modulus $\KG$, and tilt
modulus $\KTH$. The parameter $k_0$ is the spontaneous curvature and
determines the preferred lipid splay~\cite{RoLi15,Kozlov2007}.

Although the HK elastic theory assumes small deformations, Galimzyanov
{\em et al.}~\cite{C9SM02079A} have shown that energies derived from
molecular dynamics and those derived from \eqref{ansatz3} are in
agreement, even when curvatures are large.  Under spatial scales much
larger than the membrane thickness, membrane energy is
well-characterized by the Canham-Helfrich energy used throughout the
fluid-structure literature \cite{QiangDu09, Lowengrub07,KimLai2010_JCP,
Hu, HuLaiSeolEtAl2016_JCP, qua-bir2014, qua-vee-you2019}. The
Canham-Helfrich energy is actually a special case of \eqref{ansatz3}
obtained by setting ${\bf n} =  \pm {\bf N}$ (the $\pm$ depending on
orientation) and collapsing both monolayers onto the membrane midplane.

\subsubsection{Simulations of the HAP model to estimate elastic moduli
and energy}
\begin{wrapfigure}[11]{r}{0.43\textwidth}
  \centerline{\includegraphics[width=0.42\textwidth]{Figures/Flattening.pdf}}
  \caption{\label{fig:flattening} \footnotesize Example of computing
  bending modulus of a lipid bilayer from particle simulation.}
\end{wrapfigure}
We propose to {\bf (i)} determine the value of effective elastic moduli
$\KB$, $\KT$, $\KG$, and $\KTH$ and then {\bf (ii)} understand how the
model parameters $\rho,$ $\gamma$, and the particle shape map onto
elastic moduli. An accurate and robust way to measure material
properties is to track the evolution of the bilayer as it relaxes from
an initially non-equilibrium
configuration~\cite{PhysRevLett.117.188102}.
Figure~\ref{fig:flattening} shows an example where the initial
configuration of a bilayer patch containing the splay deformation
without any other components (saddle splay, twist and tilt). As the bent
membrane flattens, both the self-interaction energy $\Phi$ and the
elastic energy ${\cal W}$ decrease, and because the saddle splay, twist,
and tilt stay zero, the slope gives the bending modulus in this case.

Here we start with a particle-based bilayer in a specific
non-equilibrium shape that involves only one of the components of the
displacement in Figure~\ref{fig:deformations}, and then evolve the
particle system according to the time integration for \eqref{eq:stokes}.
Because elastic properties are independent of dissipation, we can forgo
solving for fluid velocity and set the translational velocity and
angular velocity directly proportional to the force and torque,
respectively.

This yields a dissipation of the total potential \eqref{eq:total_poten},
which stabilizes the evolving bilayer shape.  Therefore, we reconstruct
an evolving monolayer dividing surface $\Sigma$ and director field ${\bf
n}$ by interpolating the particle centers and orientations. Using
\eqref{ansatz3}, we calculate a continuum energy $\mathcal{W}$ from the
interpolated shapes.

We propose to conduct calculations similar to the example of
Figure~\ref{fig:flattening} for other elastic moduli such as the
effective twist modulus $\KT$.  Molecular dynamics investigations find a
twist modulus about 1 \kBT~\cite{LeVeWa14}, and here, the specific
non-equilibrium shape consists of a single layer of amphiphilic
particles on a hydrophobic substrate as illustrated by
Figure~\ref{fig:deformations}D. Having a nonzero twist requires nonzero
tilt because the surface gradient of the lipid director equals the
second fundamental form whenever tilt is zero locally. The twist
deformation is a fully three-dimensional deformation.
\S\ref{subsec:specific_aim_3} addresses outstanding implementation
issues like three-dimensional boundary integral equation solvers.

The gradient descent technique is ineffective for measuring the saddle
splay modulus $\KG$ because the saddle splay energy is largely invariant
under shape changes.  To evaluate $\KG$, we will combine the present
particle simulations with the string method from PI RR's work on
membrane fusion \cite{RyKlYaCo16}. The string method is a numerical
scheme that finds least energy pathways separating energy basins
\cite{doi:10.1063/1.2720838}.  In the simplified Canham-Helfrich
formulation, saddle splay energy is an exact indicator of topological
transitions, thanks to the  Gauss-Bonnet theorem \cite{TerziDeserno17}.
More generally, PI RR has shown that saddle splay acts as a topological
indicator even in the presence of nonzero tilt \cite{RyKlYaCo16}.  As a
result, saddle splay can be quantified from the transition energies of a
least energy path of pore formation (Figure~\ref{fig:saddle_splay}).

The field of membrane continuum mechanics still lacks consensus as to
whether HK energy is the appropriate functional for bilayer energy.
\textbf{(i)} Researchers have assumed that $\KT = 0$ to effect lateral
fluidity in membranes~\cite{Hamm2000, TerziDeserno17, C9SM02079A,
PhysRevE.102.042406}. A value $\KT = 0$, however, makes~\eqref{ansatz3}
a noncoercive functional. \textbf{(ii)} Recently,~\cite{TerziDeserno17}
derived a tilt curvature term that was neglected from the HK
analysis~\cite{Hamm2000}.  Later,~\cite{C9SM02079A}
and~\cite{PhysRevE.102.042406} independently identified an inconsistency
in the argument used by~\cite{TerziDeserno17} arising from a transversal
tilt invariance assumption.  In~\cite{RyKlYaCo16}, PI RR and
collaborators showed that the tilt vector leads to unphysical cusps
depending on how one accounts for membrane thickness.  \textbf{(iii)}
Theoretical analysis of lipid phase transitions predict a negative
saddle-splay modulus around $-8$ \kBT~\cite{SIEGEL2004366,
SIEGEL20085200} that gives rise to a larger energy barrier for monolayer
fusion than is found by experiments~\cite{FrRoPi17, Tran7106,
TerziDeserno17}.
\begin{wrapfigure}[11]{l}{0.32\textwidth}
  \centerline{\includegraphics[width=0.31\textwidth]{Figures/SaddleSplayDiagram.pdf}}
  \caption{\label{fig:saddle_splay} \footnotesize Example of determining
  the saddle splay modulus.}
\end{wrapfigure}

The form of the elastic energy density~\eqref{ansatz3} is the same as
the Oseen-Frank energy density for nematic liquid
crystals~\cite{ANDRIENKO2018520, Tran7106, Helfrich73}. In fact, a lipid
monolayer acts as one layer in a smectic phase~\cite{REYESMATEO1995978,
Rangamani20140463, PhysRevLett.113.248102}. Based on this observation,
we propose to examine the HK analysis to resolve the aforementioned
inconsistencies. We will expand the strain tensor in terms of a plane
perpendicular to ${\bf n}$ (instead of the monolayer tangent plane as
done in the past works) so that the gradient terms in the elastic energy
completely decouple. Using this expansion we prove an exact identity
that gives the incompressibility condition by a Steiner-type polynomial
in $\Div {\bf n}$ and $\det \nabla {\bf n}$ \cite{Fe59}. In contrast,
the works~\cite{TerziDeserno17, PhysRevE.102.042406, Hamm2000,
C9SM02079A} utilize an approximate identity for incompressibility as a
base, so we can considerably improve upon the analysis of monolayer
energy.

\subsubsection{Analysis of the HAP model in terms of the HK functional}
The HK functional~\eqref{ansatz3} has transformed how scientists understand biological membranes, yet the
theory has received little attention in terms of
mathematical analysis. In the calculus of variations, the
principal question is whether a minimizer of an energy functional
exists. In the case of~\eqref{ansatz3}, the answer is presently unknown.
With regard to the simpler Canham-Helfrich energy,~\cite{Simon1993}
proved the existence of minimizing surfaces without boundary,
and~\cite{doi:10.1137/18M1195851} proved the well-posedness of a
spatially periodic, time-dependent elastic interface problem. An
analytical challenge for the HK functional~\eqref{ansatz3} is that
surface-director coupling makes it possible to have bounded energy
monolayers with corners, and such pathological examples must be ruled
out before carrying over the arguments for the Canham-Helfrich
functional to the present setting.

Our goal here is to develop tools that can explain functionals
like~\eqref{ansatz3} from first principles in terms of HAP. The
papers~\cite{doi:10.1063/5.0009734, Seguin2012, Seguin2014} give
statistical mechanical/mean field derivations of the Canham-Helfrich
energy from a pair-potential for rod-like molecules, but do not include
tilt, which is an indispensable deformation at biological scales. To
make progress, we must first understand how the HAP functional behaves
under various limits.

We first consider HAP in the limit of vanishing screening length. As a
concrete model problem, we consider a collection of colloidal polyhedral
where the binding energy of this system can be described by a discrete,
lower semicontinuous functional $\Phi_0$ whose value is the total
surface area of the polyhedra minus two times the area of any
overlapping faces. We conjecture that the $\Phi$ energy
$\Gamma$-converges to $\Phi_0$ \cite{Mugnai2013}, meaning that any
cluster points of minimizers of $\Phi$ converge to a minimum of $\Phi_0$
in the limit $\rho \to 0$. To address this conjecture, we employ
boundary layer analysis for the screened Laplace equation~\cite{Lee2018,
Lin2015, Shibata2004,1531-3492_2006_2_357, Lee2018}. The technique of
$\Gamma$-convergence is a powerful tool for numerical approximations and
with it we can help explain unexpected phenomena like hierarchical
self-assembly in colloidal systems~\cite{Luo2019}.

\end{comment}