### 2021-11-11 NSF2021/Description/specificaim2.tex
working on SA2

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose to develop efficient high-order numerical methods to
simulate dynamic assembly of particles under HAP potential in a viscous
solvent. As described in~\eqref{eq:stokes1}--\eqref{eq:stokes3} and
illustrated in Figure~\ref{fig:flow_map}, the solvent phase is modeled
by the Stokes equations for an incompressible, viscous fluid, and these
equations are coupled to the screened-Laplace equation~\eqref{eq:SL}
through viscous and hydrophobic stress balance.

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior Dirichlet problems for the order
parameter~\eqref{eq:SL} and the Stokes
equations~\eqref{eq:stokes1}--\eqref{eq:stokes3}. If discretizing these
equations with stencil-based numerical methods, the computational domain
must be truncated, the fluid volume must be discretized, and artificial
boundary conditions must be imposed. The artificial boundary conditions
introduce additional error, and discretizing the volume results in an
excessively large system of equations. Moreover, it is difficult to
obtain high-order discretizations when the boundary surfaces are
irregular.

Alternatively, we will take advantage of the linearity of the Stokes
equations, and also apply a Newton's method to the HAP
model~\eqref{eq:SL}, where each iteration requires the solution of
linear elliptic PDEs. Then we can represent solutions of these elliptic
PDEs as layer potentials
\begin{align}
  \label{eq:LP}
  u(\mathbf{x}) = \mathcal{K}[\sigma](\mathbf{x}) = 
  \int_{\partial\Omega} K(\mathbf{x},\mathbf{y})
  \sigma(\mathbf{y})\,dS,
\end{align}
where $\sigma$ is an unknown density function, and
$K(\mathbf{x},\mathbf{y})$ is a derivative of the fundamental solution
of the PDE. By matching the prescribed boundary condition at
$\mathbf{x}_0 \in \partial\Omega$ with the limit of
equation~\eqref{eq:LP} as $\xx \rightarrow \xx_0$, a boundary integral
equation (BIE) for the density function is formed.

Numerical methods to solve BIEs have several advantages over their
PDE-based counterpart: by construction, the layer
potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(\mathbf{x},\mathbf{y})$ result in a well-conditioned linear system
that is solved with a mesh-independent number of GMRES iterations; and
high-order or spectral accuracy is attainable with appropriate
quadrature methods. The PIs have demonstrated that BIEs are a powerful
tool to simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo20}. In addition to improving
two-dimensional simulations, the present proposal extends the results to
three dimensions using a three-dimensional BIE formulation of the
screened Laplace equation~\cite{ying_2006} and a well-conditioned BIE
formulation of the mobility problem~\cite{manasthesis, rac-gre2016}.

Significant challenges to solving BIEs include solving largely stem from
solving dense linear systems. First, \S\ref{subsec:AC} describes how the
non-linear Allen-Cahn equation~\eqref{eq:SL} will be iteratively solved
using BIE formulations. Then, \S\ref{subsec:NumericalIssues} describes
how we will combine preconditioning strategies with fast summation
methods to accelerate the time to solution, and
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,

\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  J[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha (dJ^N)^{-1} F(u^N),
\end{align*}
where $dJ^N = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative
of $F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest computational challenge is
solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = J[u^N].
\end{align}
\begin{wrapfigure}[]{l}{1.8in}
  \vspace{-46pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $u_0 =
  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging since the fundamental solution is not easily
determined. We propose to instead use an inexact Newton's method where
$dJ^{-1}$ is defined in terms of the piecewise quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - u_0)^2, &\mbox{if } u \leq (u_0 + u_1)/2 \\
    (u - u_1)^2, &\mbox{if } u > (u_0 + u_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = J[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of exact and inexact Newton's method
applied to~\eqref{eq:F} by considering a finite difference method
applied to its one-dimensional version in $(0,1)$ with the Dirichlet
boundary conditions $u(0) = u_0$ and $u(1) = u_1$. These parameters
result in the double well potential illustrated in
Figure~\ref{fig:CA}(a), and both the exact and inexact Newton's methods
converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the max norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
piecewise constant approximate (blue curve).

Having demonstrated that the inexact Newton method converges in one
dimension, it will be used to solve the Allen-Cahn equation in two more
or more dimensions. Since we are interested in solving~\eqref{eq:SL} in
general unbounded geometries, we will develop an integral equation
formulation. We first partition $\Omega$ into $\Omega_0 = \{\xx \in
\Omega \:|\: u(\xx) < (u_0 + u_1)/2\}$ and $\Omega_1 = \{\xx \in \Omega
\:|\: u(\xx) > (u_0 + u_1)/2\}$, and we let $\Lambda = \{\xx \:|\:
u(\xx) = (u_0 + u_1)/2 \}$ be their common boundary points. We then let
$G_0$ be the fundamental solution of $(-\rho^2 \Delta - 2a)$ and $G_1$
be the fundamental solution of $(-\rho^2 \Delta + 2)$. Then, the
solution of~\eqref{eq:screenedPoisson} can be written as $u = u^P + u^H$
where $u^P$ is a volume potential and $u^H$ is a layer potential. The
volume potential is
\begin{align*}
  u^{P}(\xx) = \int_{\Omega_i} G_i(\xx,\yy) J[u^N](\yy)\, d\yy, 
    \quad i = 1,2,
\end{align*}
but it is discontinuous across $\Lambda$ and does not satisfy the
prescribed hydrophobic boundary conditions on the particles $\Gamma_i$.
The boundary conditions are enforced with the layer potential
\begin{align*}
  u^{H}(\xx) = \begin{cases}
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_0(\yy) \, ds_\yy + 
    \displaystyle\int_{\Gamma} \pderiv{}{\nn_\yy} G_0(\xx,\yy)
      \sigma(\yy) \,ds_\yy, \quad & \xx \in \Omega_0, \\
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_1(\yy) \, ds_\yy,
     & \xx \in \Omega_1.
  \end{cases}
\end{align*}

Then, assuming that the particles are all contained in the
hydrophobic region, a solution of the screened Laplace equation
(i.e.~equation~\eqref{eq:screenedPoisson} when $G[u^N] = 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}

% ----------------------------------------------------------------------
\subsubsection{Accelerating the iterative solvers}
\label{subsec:NumericalIssues}

%
%\begin{wrapfigure}[20]{r}{2.5in}
%\vspace*{-13pt}
%\centerline{\includegraphics[width=2.4in]{figures/integrands}}
%\vspace*{-13pt}
%  \caption{{\footnotesize The Fourier modes of
%  $K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ where
%  $|\mathbf{x_0}| = 0.99$, $\mathbf{y} \in \partial\Omega$, and
%  $\Omega$ is the unit disk. The trapezoid rule applied to the red
%  integrand has large amounts of error, but the trapezoid rule is much
%  more accurate when applied to the blue integrand. Since the difference
%  between the red and blue curves can be accurately computed with the
%  barycentric quadrature rule, the quadrature error applied to layer
%  potentials for the screened Laplace equation will be uniformly bounded
%  in $\Omega$.}}
%\label{fig:integrands}
%\end{wrapfigure}
Discretizations of carefully chosen BIEs can be solved with a
mesh-independent number of GMRES
iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required CPU
time is proportional to the cost of a matrix-vector multiplication that
can be performed in optimal or near-optimal time with the fast multipole
method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2, fmm3,
fmm4, fmm6, fmm7, fmm8}. PI BQ is experienced with applying FMMs for the
Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the screened
Laplace equation~\cite{kro-qua2011, qua2011}.  Alternatively, fast
direct solvers avoid iteration~\cite{fds2, fds3, fds4, fds5, fds6, fds7,
ho2016cpam1, minden2016, minden2017siammms}, but often require a large
amount of computational overhead, and this makes them less practical for
problems with moving geometries. Alternatively, fast direct solver
techniques can be used to develop efficient preconditioning strategies
such as the inverse Fast Multipole Method (IFMM)~\cite{cou-pou-dar2017}.
PI BQ used the IFMM to precondition Stokes
equations~\cite{qua-cou-dar2018}. A suite of other preconditioners
including sparse approximate inverses~\cite{che2000} and
multigrid~\cite{hem-sch1981, sch1982} will be investigated.
%
%Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
%particles, and this results in nearly-singular integrands. Quadrature
%methods for these integrands have received a lot of attention in two and
%three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
%davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
%khayat_2005, schwab_1992, ying_2006, beale1, schwab_1992, ggq1, ggq2,
%ggq3, helsing_2008a, helsing_tutorial_2012, klockner2013jcp, qbx2,
%wala2019jcp, af2018sisc, siegel2018jcp, rachh2017jcp, ding2019arxiv,
%bar2014}. The trapezoid rule is the workhorse for two-dimensional BIEs
%since it achieves spectral accuracy when integrands are not
%nearly-singular~\cite{tre-wei2014}. To address nearly-singular
%integrands of two-dimensional BIEs, we will use a {\em barycentric
%quadrature rule} that only requires a slight modification of the
%trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this method
%requires the layer potential to satisfy Laplace or Stokes
%equations~\cite{bar-wu-vee2015, chi-moo-qua2020}, but the PIs will
%extend the method to layer potentials for the screened Laplace equation.
%This will be done by recognizing that the fundamental solution of the
%screened Laplace equation can be decomposed as $K(\mathbf{x},\mathbf{y})
%= K_1(\mathbf{x},\mathbf{y}) + K_2(\mathbf{x},\mathbf{y})$, where
%$K_1(\mathbf{x},\mathbf{y}) = -\log|\mathbf{x} - \mathbf{y}|$ and
%$K_2(\mathbf{x},\mathbf{y}) = K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x}
%- \mathbf{y}|$. Then, layer potentials involving
%$K_1(\mathbf{x},\mathbf{y})$ can be accurately computed with the
%barycentric quadrature rule~\cite{ioa-pap-per1991}, and layer potentials
%involving $K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with
%the trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
%$\mathbf{y}$. The proposed method is demonstrated in
%Figure~\ref{fig:integrands}, where the amplitude of the Fourier modes of
%$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
%plotted. The quadrature error of the trapezoid rule is the sum of the
%unresolved Fourier modes.

% ----------------------------------------------------------------------
\subsubsection{Eliminating contact with adaptive time stepping and
repulsion}
\label{subsec:timeStepping}

A numerical issue when simulating the self-assembly of amphiphilic
particles is avoiding particle collision. The hydrophobic attraction
potential drives the amphiphilic particles towards one another and
this leads to physical contact in
finite time. Such particle collisions in a dense, rigid-body suspension
are a great challenge and can be a bottleneck in large-scale simulations.
We propose two algorithms to remedy this computational challenge:
high-order adaptive time stepping and repulsion forces.

PI BQ developed a high-order adaptive time stepping method for
hydrodynamic suspensions \cite{qua-bir2016}, and it has served as a
robust method to simulate processes including mixing and adhesion in
suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}. The method uses
a single-step, high-order time stepping method and a computationally
cheap estimate of the error. The proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order, single-step method to achieve high-order accuracy.
To estimate the error, at each time step, we will compute the total force
and torque of the system which is computationally cheap and physically
zero. PI BQ's experience is that error estimates based on physical
constraints such as force- and torque-free conditions appropriately
determine how the time step should be adjusted so that the dynamics are
resolved without the computational expense of techniques such as
embedded Runge-Kutta methods, step-doubling, and Richardson
extrapolation.

%%
%\begin{wrapfigure}[13]{r}{0.30\textwidth}
%\centerline{\includegraphics[width=0.29\textwidth]{figures/BIError.pdf}}
%  \vspace{-8pt}
%\caption{
%\label{fig:bierror}  
%\footnotesize The false color map shows how numerical quadrature of
%  layer potentials loses accuracy near the particle boundaries.  The
%  color bar is for $\log_{10}.$}
%\end{wrapfigure}
%%
Our previous work avoided particle contact through Lennard-Jones body
forces~\cite{Fu2018_SIAM}. Such steep, short-range, steric interactions 
introduce numerical stiffness and limit the time step size. To
remove this stiffness, we propose a geometry-based contact
method~\cite{har-pon-sor-zor2011}. This method has been applied to
vesicle suspensions~\cite{lu-rah-zor2017} and to rigid body suspensions in
two dimensions~\cite{bys-sha-qua2020} and three
dimensions~\cite{Yan2019}. These methods determine repulsion force by
solving a non-linear complementarity problem with a geometric constraint
that the configuration is non-overlapping. In this manner, contact is
avoided without the excessively small time steps required by stiff
Lennard-Jones forces.

%%-------------------------------------------------------------------------------
%\subsubsection{Novel reciprocal identities}
%\label{subsec:reciprocal}
%The force and torque formulas~\eqref{forceandtorque} use the
%hydrophobic stress along the particle boundaries. Unfortunately,
%layer potentials for $\nabla u$ involve hyper-singular integrals and
%this introduces large amounts of quadrature error on the particle
%boundaries (Figure~\ref{fig:bierror}). Therefore, it is useful to devise
%reciprocal identities for the force and torque on body $i$ that do not
%require integration along body $i$. One identity, that we have proved,
%is
%\begin{align}
%    \label{eq:reciprocal}
%{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
%{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
%  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
%  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
%\end{align}
%for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
%equation when only the contribution from particle $i$ is considered and
%$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j \mathbf{I} +
%\rho\left((\nabla u_i \cdot \nabla u_j) \mathbf{I} - 2 \nabla u_i
%\otimes \nabla u_j\right)$. Equation~\eqref{eq:reciprocal} does not
%require knowledge of $\nabla u$ on particle boundary $i$, and this is
%enormously beneficial for calculating the hydrophobic force and torque.
%
%% ----------------------------------------------------------------------
%\subsubsection{Fluctuating hydrodynamics}
%\label{subsec:fluctuating}
%Once algorithms that address quadrature, adaptive time stepping, and
%contact are implemented, we will embark on incorporating fluctuating
%hydrodynamics that are especially important at scales of amphiphilic
%particles. Bao et al.~\cite{Bao17,Bao18} developed two-dimensional BIE
%formulations for Brownian rigid bodies. To satisfy the covariance of the
%rigid body velocities, they use an ill-conditioned, first-kind integral
%equation method. The proposed work will consider a layer potential
%formulation that has much better conditioning properties. This strategy
%was attempted in~\cite{Bao18}, but they claim that a necessary
%regularization of the covariance matrix leads to a drastic loss of
%accuracy in numerical fluctuation-dissipation balance. However, only one
%regularization strategy was considered, and as PI BQ has shown in
%previous work, the regularization choice significantly affects the
%resulting physics~\cite{ong-chr-qua2017}. Therefore, different
%regularization strategies will be investigated with the goal of finding
%well-conditioned BIE formulations that control the accuracy of the
%fluctuation-dissipation balance.