### 2022-10-11 doc/PRF2022/Reviews/ResponseToReviewer1.tex
adding reviews

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{color}
\usepackage{newtxmath}
\newcommand{\comment}[1]{{\color{blue} #1}}

\begin{document}

\noindent
Dear Prof.~Cecile Cottin-Bizonne,
\\ \\
\noindent
Thank you for handling the manuscript, and we would like to thank the
reviewer for their positive and constructive comments. The attached
manuscript addresses all the reviewer's comments and the result is what
we believe is a stronger manuscript. An itemized list of the changes
addressing the reviewer's comments are below. Changes to the manuscript
from Reviewers 1 and 2 are colored in \textcolor{red}{red} and 
\textcolor{magenta}{magenta}, respectively. \\ \\
\noindent
Sincerely,
\\
\noindent
Szu-Pei Fu, Rolf Ryham, Bryan Quaife, and Yuan-Nan Young

\section*{Response to Reviewer 1}
\noindent
\comment{The manuscript discusses the self-assembly of amphiphilic Janus
particles and their dynamics in various flow conditions. Different
morphologies are recovered for different boundary conditions which are
developed by varying the hydrophobicity on Janus particle surface.
These various boundary conditions give rise to unilamellar,
multilamellar and striated structures. Finally, these structures are
subjected to linear shear flow and Taylor-Green flow. The subject of
study is interesting to soft matter community.}

\begin{itemize}
  \item {\bf Response to Summary}: 	
  \begin{quotation}
%    \noindent
%    Can put verbatim change in this environment.
  \end{quotation}
\end{itemize}

\noindent
\comment{My main criticism is
the following: 1. Unilamellar, multilamellar and striated structured are subjected to
various flow conditions, with increasing shear rate or strength of the
flow (for TG case). For all structures in both flows, the main result
is that these structures break eventually either for increased shear
rate or increased flow rate. But this is expected. What do all these
results mean? I feel that a constructive message is missing.}

\begin{itemize}
  \item Response:

  \item Change:

\end{itemize}

\noindent
\comment{2. And the discussion section is not written well. Please write it
properly.}

\begin{itemize}
  \item Response:

  \item Change:

\end{itemize}

\noindent
\comment{Minor comments: 1. Page 18: The section ends with 
the statement “In previous work, Fu
et al. [18] determined the permeability constant of particle-based
vesicles in the context of shear background flow [53, 54]. “. I find
the statement abrupt and not related to the earlier text.}

\begin{itemize}
  \item Response:

\item Change:

\end{itemize}

\noindent
\comment{2. Fig. 8: Please describe what the arrows mean in (a), (b), (c) and
(d).}

\begin{itemize}
  \item Response:

  \item Change:

\end{itemize}

\end{document}