### 2022-11-14 NSF2022/SupplementaryDocs/suggestedreviewers.tex

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{enumitem}
\pagestyle{empty}
\begin{document}
\topmargin0.25in
%\thispagestyle{empty}
\sloppy

\noindent
\today \\ \\
Dear Program Officer:
\\ \\
\noindent
The following individuals have the expertise to fairly evaluate the 
proposal entitled ``Collaborative Proposal: Self-organization modeled by
moving domain elliptic PDEs". They are organized according to specific
themes of the proposed research.
\\ \\
Sincerely, \\ \\
Bryan Quaife, Rolf Ryham, and Yuan-Nan Young \\

\subsubsection*{Fluid structure interactions}
\begin{tabular}{ll}
  Suncica Canic & canics@berkeley.edu \\
  Shibin Dai & sdai4@ua.edu \\
  Giovanna Guidoboni \qquad & guidobonig@missouri.edu \\
  Keith Promislow & kpromisl@math.msu.edu
\end{tabular}

\subsubsection*{Integral Equation Methods}
\begin{tabular}{ll}
  Alex Barnett & abarnett@flatironinstitute.org \\
  Andrew Christlieb \qquad &christli@msu.edu \\
  Tony Gao & gaotong@egr.msu.edu \\
  Shuwang Li & sli@math.iit.edu \\
  John Lowengrub & jlowengr@uci.edu
\end{tabular}

\subsubsection*{Fluid Mechanics and Soft Matter}
\begin{tabular}{ll}
Silas Alben & alben@umich.edu \\
Robert Krasny & krasny@umich.edu \\
Saverio Spagnolie \qquad & spagnolie@math.wisc.edu \\
Petial Vlahovska & petia.vlahovska@northwestern.edu \\
Noel Walkington & noelw@andrew.cmu.edu 
\end{tabular}

\subsubsection*{Amphiphiles}
\begin{tabular}{ll}
Andrea Bertozzi & bertozzi@math.ucla.edu \\
Bob Guy & guy@math.ucdavis.edu \\
Alice McGillman \qquad &  \\
Mike Shelley & mshelley@flatironinstitute.org 
\end{tabular}

\end{document}