### 2022-11-14 NSF2022/SupplementaryDocs/excludedreviewers.tex

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{enumitem}
\pagestyle{empty}
\begin{document}
\topmargin0.25in
%\thispagestyle{empty}
\sloppy

\noindent
\today \\ \\
Dear Program Officer:
\\ \\
Professors Alex Donev (donev@courant.nyu.edu) and Yoichiro Mori
(y1mori@sas.upenn.edu) are competitors that are working on similar
problems to those described in the proposal entitled ``Collaborative
Proposal: Self-organization modeled by moving domain elliptic PDEs". We
request that they are not chosen to review the proposal.
\\ \\
Sincerely, \\ \\
Bryan Quaife, Rolf Ryham, and Yuan-Nan Young

\end{document}