### 2022-11-15 NSF2022/Description/priorsupport.tex

\section{Relevant Results from Prior NSF Support}
\noindent {\bf Rolf Ryham}: no prior NSF support.

\noindent
{\bf Yuan-Nan Young}: {\em NSF-DMS-1222550, Mathematical and
experimental study of lipid bilayer shape and dynamics mediated by
surfactants and proteins}, \$212,603, 9/15/2012 - 08/31/2016 (with
no-cost extension), PI. {\em Intellectual merit:} The focus of this
grant is modeling the interaction between a pure lipid bilayer membrane
with surfactant, cholesterol, and protein.

\noindent
{\it Broader impacts:} 
One PhD student (Szu-Pei Fu) was funded to work with PI Young, and work
has resulted in seven papers~\cite{Nganguia2013_PoF, Nganguia2013_PRE,
Young2014_JFM, Young2015_PoF, Nganguia2015_CiCP, Pak2015_PNAS,
fu2015pre}. 
%PI Young has been actively involved with promotion of
%underrepresented students at NJIT. One PhD student, Herve Nganguia, is
%now an Assistant Professor at Towson University. PI Young has taught a
%broad spectrum of courses in fluid mechanics and applied math modeling.

\noindent
{\bf Bryan Quaife}: {\em NSF DMS-2012560, Erosion, Transport, and
Dispersion in Granular and Porous Media}, \$249,636,
08/01/2020--07/31/2023, PI. {\em Intellectual Merit:} The goal of this
research is to develop high-order numerical methods to
simulate hydrological processes including erosion.

\noindent
{\it Broader impacts:} 
A third-year PhD student (Jake Cherry) has been assigned to this
project. One paper has been accepted~\cite{che-lin-her-qua2022} and
another is under review~\cite{moo-che-chi-qua2022}.

\section{Project Management, Collaboration Plan, and Schedules of
Research Tasks}
\setlength{\parindent}{0pt}
% N.B. wrapfigure is not compatible with \noindent.

\begin{wrapfigure}[9]{r}{0.4\textwidth}
  \vspace{-5pt}
  \includegraphics[width=0.4\textwidth]{figures/gantt.jpg}
  \vspace{-20pt}
  \caption{\label{fig:schedule} \footnotesize Schedule for the proposed
  work, measured in quarters from the beginning of the project. The lead
  PI of each Specific Aim will be Ryham (red), Quaife (blue), and Young
  (green).}
\end{wrapfigure}

\begin{comment}
\begin{wrapfigure}[11]{r}{0.56\textwidth}
\vspace{-19pt}
%\definecolor{barblue}{RGB}{51,102,254}
%\renewcommand\sfdefault{phv}
%\renewcommand\mddefault{mc}
%\renewcommand\bfdefault{bc}
%\sffamily
\begin{ganttchart}[
    canvas/.append style={fill=none, draw=black!5, line width=.75pt},
        x unit =4.5mm,
        y unit chart =\baselineskip,
    hgrid style/.style={draw=black!5, line width=.75pt},
    vgrid={*1{draw=black!5, line width=.75pt}},
    title/.style={draw=none, fill=none},
    title label font=\bfseries\footnotesize, % numbers across the top
    title label node/.append style={below=3pt},
    include title in canvas=false,
    bar label font=\mdseries\footnotesize\color{black!70}, 
    % titles across the left
    %bar label node/.append style={left=2cm},
    bar/.append style={draw=none, fill=blue},
    bar incomplete/.append style={fill=blue},
    bar progress label font=\mdseries\footnotesize\color{black!70},
    milestone label font=\mdseries\small\color{black!70},
        milestone left shift =0.9,
        milestone right shift =0.1,
        % Don't draw group bars
        group height =0,
        group peaks height =0,
        group label font =\bfseries\small,
]{1}{12}
\gantttitle[
    title label node/.append style={below left=3pt and -6pt}
]{QUARTERS:\quad1}{1}
\gantttitlelist{2,...,12}{1} \\
%\ganttgroup{Tasks\hfill}{1}{12}\\
\ganttbar[bar/.append style={fill=red}]{Specific Aim 1}{1}{12} \\
\ganttbar[bar/.append style={fill=blue}]{Specific Aim 2}{1}{12}\\
\ganttbar[bar/.append style={fill=green}]{Specific Aim 3}{1}{12}\\
\ganttbar[bar/.append style={fill=black}]{Program Management}{1}{12}
\ganttbar[bar/.append style={fill=magenta}]{}{4}{4}
\ganttbar[bar/.append style={fill=magenta}]{}{8}{8}
\ganttbar[bar/.append style={fill=magenta}]{}{12}{12}
\end{ganttchart}
\vspace{-10pt}
\caption{Schedule for the proposed work, measured in quarters from the
  beginning of the project. The lead PI of each Specific Aim will be
  Ryham (red), Quaife (blue), and Young (green). Project management will
  be continuous (black), but we anticipate extra attention will be
  necessary to align objectives in quarters 4, 8, and 12 (magenta).}
\label{fig:schedule}
\end{wrapfigure}
\end{comment}
\textbf{Project management}: 
%
The success of the proposed research requires complementary expertise
and collaborative efforts in physics, applied mathematics, and
scientific computing. PI Ryham has been working on mathematical modeling
with a strong analytical background. PI Young has been working on many
areas of fluid dynamics with applications to soft matters. 
PI Quaife has been working on integral equation methods,
fast algorithms, and their applications to fluid dynamics.  
Their recent collaborative works on the granule-based vesicles
in two dimensions~\cite{FuQuRyYo22, fu-ryh-qua-you2022} has provided a
solid foundation for the proposed research.

\smallskip

\textbf{Collaboration plan}: 
%
PI Quaife is responsible for the overall management of the project.
The research work is structured to meet the tasks
in \S\ref{sec:proposed-work}. The PIs and students will meet
on Zoom and in person when possible. PI Ryham and PI Young
will conduct biweekly meetings in person, with PI Quaife joining over
Zoom from Florida. The PIs will share software packages, paper sources,
and references on a common \textsf{Git} repository. The resulting
software packages will be posted on the \textsf{Github} software
repository.
%
%\smallskip
%
\textbf{Research Schedule:} The detailed schedule for the proposed work
is shown in Figure~\ref{fig:schedule}. Each PI will spearhead a
Specific Aim, and program management will happen
throughout, with special coordination taken in quarters 4, 8, and
12.