### 2022-11-15 NSF2022/Description/broaderimpacts.tex
ed imp done

\section{Broader Impacts}
\label{sec:BroaderImpacts}
This project aims to advance the mathematical modeling of collective
dynamics of amphiphilic granules. The framework uses a new PDE-based
formulation that accounts for important and complex systems in soft
matter. These complex systems include optimal shape design in
metamaterials and fusion and fission of amphiphilic bilayer membranes.
The simulations will be performed with computational tools designed to
solve the governing PDEs both efficiently and with high-order accuracy.
The models describing granular systems could be transformative in
biomedicine and material science. The research draws from expertise in
scientific computing, physics of fluids, and mathematics. The
mathematical component incorporates leading variational techniques and
offers insight into fundamentals of self-organization and collective
dynamics. The project brings socially consequential research into the
classroom and offers undergraduates the opportunity to train alongside
faculty and graduate students. With its unique combination of
mathematical modeling, analysis, and scientific computing, the project
highlights the potential advancement of STEM from applied mathematics.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
To foster training in mathematical sciences,
the proposal supports two undergraduate researchers
(URs) per year at Fordham University---for eight summer weeks.
The UR will collaborate directly with PI Ryham,
receive tutorials e.g., numerical methods,
and be guided on mathematical writing and presentation.
Fordham Mathematics recently completed a collaboration studio 
purpose-built for UR support.
Work products include a summary manuscript,
a presentation in the Fordham Undergraduate Research Symposium
and departmental seminars, and participation in a national conference.
We have outlined three projects for the present proposal.
\begin{enumerate}[noitemsep,topsep=0pt]
\item Translate project code from MATLAB into a compiled language 
\item Use machine learning to post-process data of granule dynamics
\item Investigate geometrical self-assembly and generate visualizations
\end{enumerate}
PI Ryham has worked side-by-side with dozens of
undergraduates and high-schoolers in
supported summer research, Girls in Science Technology and Mathematics,
and the Clare Boothe Luce Scholarship program
and the activity benefited their careers in
machine learning, water resources engineering, threat detection,
and the NYC Department of Education.
Presently PI Ryham is mentoring a Clare Boothe Luce Scholar on election redistricting.
The PI is finishing works on the energy density lipid  bilayers
with a recently graduated UR, now working at Raytheon Technologies.
His published works~\cite{RYHAM20112929, RyWaCo13, RyKlYaCo16} include
four undegraduate coauthors.

To increase the participation of women,
persons with disabilities and underrepresented minorities in STEM,
the PIs will specifically target students
whose socio-economic background prevents them from participating in
out-of-state research experiences.
The inclusion objective will be aided by the fact that  
Fordham's class of 2026 weighs in with 63 \% women
and more than 46 \% domestic students of color. 
PI Ryham will solicit applicants from
Fordham University's Collegiate Science and Technology Entry Program.
Students from the Bronx High School of Science (PI Ryham) and Newark
Science Park High School (PI Young) will also be encouraged to join the
research team.

PI Young has been actively involved with promotion of
underrepresented students at NJIT. One PhD student, Herve Nganguia, is
now an Assistant Professor at Towson University. PI Young has taught a
broad spectrum of courses in fluid mechanics and applied math modeling.
PI Quaife will work with undergraduates and high school
students through the Undergraduate Research Opportunity Program and
Young Scholar's Program, as he has done in the past.

%To maximize vertical integration, the PIs and personnel will travel to
%New York for one or two weeks during the summers. 
Finally, we will
create several modules to include in our courses. These modules will
introduce topics from numerical linear algebra and optimization and
highlight concrete applications of mathematics based on the ideas of the
proposal.

\section{Intellectual Merit}
\label{sec:IntellectualMerit}
This collaborative project focuses on modeling, simulation, and analysis
of self-assembly and collective dynamics of a suspension of granules with 
non-uniform hydrophobicity on their boundaries. The main
ingredient is a nonlocal interaction through the solution of moving
domain elliptic PDEs that encompasses long-range (non-additive) amphiphilic and
short-range steric interactions. The PIs have validated this
coarse-grained model against well-studied vesicle hydrodynamics. The
technical research tasks include quantifying collective properties of
ensembles of granules with tunable hydrophobicity, analyzing the mathematical model in 
distinguished limits to make connection with continuum models,
designing efficient and
high-order numerical algorithms for large-scale two- and
three-dimensional simulations with confinement, and developing a kinetic
theory to quantify the both the polar and nematic characteristics in the collective dynamics.
%
%The purpose of this research is to reach interesting physical phenomena
%with less computational cost than molecular dynamics, and account for
%more general features that continuum theory misses. The main ingredient
%is defining a nonlocal interaction through the solution of an elliptic
%boundary value problem that has the phenomenological characteristics of
%long-range hydrophobic attraction. This minimal model, while intuitive,
%is quite a general description of amphiphiles in solvent and gives rise
%to rich phenomena from Janus particle aggregates to correctly predicting
%elastic properties of bilayer. The technical research tasks include
%quantifying collective properties of amphiphilic ensembles, improving on
%mathematical models, efficient, high-order numerical algorithms for
%large-scale simulations, and incorporating external fields through
%electric charge.