### 2023-03-10 doc/PRF2022/PRF_main.tex
Results and SM updated

%% ***A*** Start of file apstemplate.tex ****** %
%%
%%
%%   This file is part of the APS files in the REVTeX 4.2 distribution.
%%   Version 4.2a of REVTeX, January, 2015
%%
%%
%%   Copyright (c) 2015 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.2
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showkeys' option to make keywords appear
%\documentclass[aps,prl,preprint,groupedaddress]{revtex4-2}
\documentclass[prb,preprint,showpacs,preprintnumbers,amsmath,amssymb,longbibliography]{revtex4-1}

%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-2}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-2}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
\bibliographystyle{apsrev4-2}

%\usepackage{showkeys}
\usepackage{todonotes}
\usepackage{graphicx}
%\usepackage{epstopdf,epsfig}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{tabularx}
\usepackage{color,soul}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newcommand{\RomanNumeralCaps}[1]
\linenumbers

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\dd}{\mathbf{d}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\DDD}{\boldsymbol{\mathcal{D}}}
\newcommand{\SSS}{\mathcal{S}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\JJ}{\mathbf{J}}
\newcommand{\kk}{\mathbf{k}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ttau}{\boldsymbol{\tau}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\ee}{\mathbf{e}}
\newcommand{\ff}{\mathbf{f}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\zz}{\mathbf{z}}
\newcommand{\XX}{\mathbf{X}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}
\renewcommand{\tt}{\mathbf{t}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\jump}[1]{[\![ #1 ]\!]}
\newcommand{\kbt}{\mathrm{k}_{\text{B}}\mathrm{T}}
\newcommand{\KBT}{k$_{\text{B}}$T}
\newcommand{\kb}{\kappa_{\text{B}}}
\newcommand{\ka}{\kappa_{\text{A}}}

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Effects of Tunable Hydrophobicity on the Collective Hydrodynamics of Janus Particles under Flows}

% repeat the \author .. \affiliation  etc. as needed
% \email \thanks, \homepage, \altaffiliation all apply to the current
% author. Explanatory text should go in the []'s, actual e-mail
% address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro foreach each type of information

% \affiliation command applies to all authors since the last
% \affiliation command. The \affiliation command should follow the
% other information
% \affiliation can be followed by \email, \homepage, \thanks as well.
\author{Szu-Pei Fu$^{1}$, Rolf Ryham$^{2}$, Bryan Quaife$^{3}$ and Y.-N. Young$^{4}$}
%\email{peter.fu@trincoll.edu}
\affiliation{$^1$Department of Mathematics, Trinity College, Hartford, Connecticut 06106, USA\\
$^2$Department of Mathematics, Fordham University, Bronx, New York 10458, USA \\
$^3$Department of Scientific Computing, Florida State University, Tallahassee, Florida 32306, USA\\
$^4$Department of Mathematical Sciences, New Jersey Institute of Technology, Newark, New Jersey 07102, USA}

%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
%\author{Rolf Ryham}
%\email{rryham@fordham.edu}
%\affiliation{Department of Mathematics, Fordham University, Bronx, New York 10458, USA}
%
%\author{Bryan Quaife}
%\email{bquaife@fsu.edu}
 %\affiliation{Department of Scientific Computing, Florida State University, Tallahassee, Florida 32306, USA}
 %\author{Y.-N. Young}
 %\email{yyoung@njit.edu}
 %\affiliation{Department of Mathematical Sciences, New Jersey Institute of Technology, Newark, New Jersey 07102, USA}

%Collaboration name if desired (requires use of superscriptaddress
%option in \documentclass). \noaffiliation is required (may also be
%used with the \author command).
%\collaboration can be followed by \email, \homepage, \thanks as well.
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in biology. To understand how
  hydrodynamic flow may be used to actively control self-assembly of
  Janus particles (JPs), we use a model recently developed for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  background flow (JFM, 941, 2022). We investigate how various
  morphologies arise from tuning the hydrophobic distribution of the
  JP-solvent interface. We find JPs assembled into uni-lamella,
  multi-lamella and striated structures. To introduce dynamics, we
  include a linear shear flow and a steady Taylor-Green mixing flow, and
  measure the collective dynamics of JP particles in terms of their (a)
  free energy from the hydrophobic interactions between the JPs, (b)
  order parameter for the ordering of JPs in terms of alignment of their
  directors, and (c) strain parameter that captures the deformation in
  the assembly. We characterize the effective material properties of the
  JP structures and find that the uni-lamellar structures increases
  orientation order under shear flow, the multilamellar structure
  behaves as a shear thinning fluid, and the striated structure
  possesses a yield stress. These numerical results provide insights
  into dynamic control of non-equilibrium active biological systems with
  similar self-organization.
\end{abstract}

%%%

% insert suggested keywords - APS authors don't need to do this
%\keywords{}

%\maketitle must follow title, authors, abstract, and keywords
\maketitle

% body of paper here - Use proper section commands
% References should be done using the \cite, \ref, and \label commands

\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}
Janus particles are colloids combining two dissimilar chemical or
physical functionalities at their opposite
sides~\cite{KirillovaMarschelkeSynytska2019}. Self-propelling Janus
particles, for example, with a permanent biphasic asymmetry, have
emerged as a rich chemical platform for the exploration of active
matter~\cite{Meredithetal2022}.
In the absence of mobility and any
imposed flow, Janus particles (JPs) self-assemble
into oligomers of various geometries
and sizes
depending on the interactions between JPs and the 
viscous solvent~\cite{Bradley2017, KangHonciuc2018,
HongCacciutoLuijtenGranick2008},
with tunable functions for biomedical engineering
applications~\cite{GheisariSahfieeAbbasiEtAl2021_DMR,
LiuYangHuangEtAl2016_Angew, LiWangYaoEtAl2019_Nanoscale,
Bradley2016,Zarzaretal2015, KirillovaMarschelkeSynytska2019}.

When driven by an external flow, dynamic rearrangement of JPs emerges
naturally from the interactions between particles and fluid in colloidal
matter~\cite{RevModPhys.93.025008}. Such colloidal hydrodynamics
belong to a wide class of nonequilibrium self-organization in physics,
often with complexity and features similar to that of biological systems
such as living cells, bacterial baths, and animal
flocks~\cite{CollardGrosjeanVandewalle2020, Vutukuri2020}. A
long-standing challenge in fluid mechanics and material science is to
solve the ``inverse" problem of creating a model colloidal system that
will self-assemble into a prescribed
structure~\cite{PhysRevLett.128.256102}. For example, the framework of
geometrical frustration, which has been used to explain disordered
systems, could instead be used to design new, ordered systems, through
specific choices for the shapes or interactions of the
particles~\cite{Manoharan2015_Science}.

The surface of a Janus particle may have a dissimilar spatial
distribution of hydrophobicity~\cite{doi:10.1021/la503455h}. The
many-body hydrodynamics of amphiphilic Janus particles assembled as
vesicles suspended in a viscous fluid in the inertialess regime (zero
Reynolds number) has been studied using boundary integral numerical
simulations~\cite{Fu20,Fu2022_JFM}. The dynamics of the JP suspension
results from the combination of long-range hydrodynamic interaction and
non-local interactions between JPs through the distribution of a
hydrophobic attraction potential (HAP).
%
In a quiescent flow, numerical simulations of a JP suspension showed
self-assembly into micelles and bilayers of JPs that provide an
alternative means for computing the mechanical moduli of a colloidal
membrane~\cite{NaTr00, Fu20, KrFiGuKaHa13}. Under background flows, the
hydrodynamics of a JP vesicle (a self-enclosed bilayer of JPs) exhibit
many familiar behaviors of a vesicle: elongation and alignment along the
extensional direction, tank-treading, and rupture of a vesicle under
shear flow~\cite{Fu2022_JFM, grandmaison_brancherie_salsac_2021,
D2SM00179A,keller_skalak_1982, Finken08, Shaqfeh11}.

Molecular dynamics (MD) and Monte Carlo provide another avenue for
simulating the interaction between JPs, solvents, and
substrates~\cite{Brandner2019, Baniketal2021,
HongCacciutoLuijtenGranick2008, C9NR05885K}. These methods often use
pair-potentials to describe the interaction between JPs, for example by
prescribing angle dependent forces and torques that bring the
hydrophobic sides of two amphiphilic JPs into opposition. The HAP
formulation of the present work also has attractive, long-range forces
and torques, but they are instead derived from a boundary value problem
for the molecular structure of water~\cite{Ma77, GoHaKo94, ErLjCl89,
Lietal05, Israelachvili80}. Unlike in MD or Monte Carlo simulations, the
HAP interactions are nonadditive~\cite{Fu20}, so that the interactions
between a pair of JPs is affected by the presence of other particles.

\hl{Qi Wang has a preprint "A hybrid phase field method for fluid-structure interactions in viscous fluids"}.

In this work, we take advantage of the flexibility of the HAP model to
examine the effects of varying the distribution of hydrophobicity on JP
surfaces. Such variation of the boundary condition on JP surfaces has
been realized experimentally by using chemicals to adjust the polarity
of the viscous solvent~\cite{Zarzaretal2015,
KirillovaMarschelkeSynytska2019, doi:10.1021/la503455h}. We show that a
simple tuning of the hydrophobic distribution leads to transitions from
unilamellar to multilamellar or striated superstructures of JPs.
Focusing on the fluid-structure interactions that correspond to such
transitions, we investigate the deformation of these novel structures in
background flows and map out their collective behavior away from
equilibrium. Looking forward, including other fields like electric
potential for JPs synthesized with charged
polymers~\cite{HongCacciutoLuijtenGranick2008, doi:10.1021/la503455h,
KangHonciuc2018}, is straightforward within the context of the boundary
integral representations~\cite{kohl-cor-che-vee22}, opening further
lines of investigation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations: Hydrophobic Attraction Potential Mobility Problem} 
The governing equations are a system for the position and
orientation of a collection of rigid JPs~\cite{Fu2022_JFM}.
We first pose the
Stokes equations for the mobility problem giving the hydrodynamic
interactions for the particle suspension. The hydrophobic forces come
from solving a screened Laplace equation. Particle collisions are
avoided through a near-field, pair potential.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mobility problem}
%A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
%RJR: In the above sentence, suspection and immersed are redundant 
The JPs are disks of radius $c$ suspended in a viscous solvent
with center $\aa_i$, and orientation $\theta_i$
relative to the horizontal axis, where $i = 1, \ldots, N_b$ and $N_b$ is
the number of particles. The domain $\Omega \subset \mathbb{R}^2$ is
the solvent phase and $t$ is time. The boundary of $\Omega$ is
$\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where $\Gamma_i$
is the boundary of Janus particle $i$.
The sets $\Omega$, $\Gamma_1$, \dots, and $\Gamma_{N_b}$ depend on $t$.
Assuming inertial terms are negligible, we have the Stokes equations
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
\label{eq:stokesMass}
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
\label{eq:stokesFarfield}
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
solvent viscosity. The solvent velocity satisfies the no-slip boundary
condition for a rigid body motion
\begin{align}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.

To obtain forces, define the free
energy~\cite{Fu20, Fu2022_JFM}
\begin{align}
\label{eq:free_energy}
F = \gamma
  \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} u^2 \right)
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right),
\end{align}
where $u(\xx,t)$ is as the order parameter for water\cite{Ma77, GoHaKo94},
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
The order parameter $u(\xx,t)$ is assumed to minimize the free energy
$F$ and therefore satisfies the screened Laplace equation
boundary value problem
\begin{gather}
  \label{eq:SL}
  -\rho^2 \Delta u + u = 0,\quad \xx \in \Omega, \\
  \label{eq:SLbc}
  u = g, \quad \xx \in \bd\Omega, \quad 
  u \rightarrow 0, \quad |\xx| \rightarrow \infty.
\end{gather}
The boundary condition $g$ encodes hydrophobic properties of the
particle-solvent interface.
The dimensionless repulsion profile $P$ takes the form $P(s) = 1 -
\sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$. The parameter $\rho_0$ is the distance 
below which steric repulsion becomes important and 
$M$ is the repulsion modulus.

Letting $\nnu$ be the particle outward normal,
the force $\FF_i$ and torque $T_i$ coming from the free energy
acting on $\Gamma_i$ are~\cite{Fu20}
%from the
%free energy~\cite{Fu20}, we compute the variation of the domain of the
%free energy~\eqref{eq:free_energy} subject
%to~\eqref{eq:SL}--\eqref{eq:SLbc},
%and obtain 
\begin{align}
  \label{eq:FandTdef}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{align}
Here,
\begin{align}
\mathbf{T} = \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right].
\end{align}
is the second-order hydrophobic stress tensor.
Repulsion is rotationally symmetric and
does not enter $T_i$.

Without inertia, $(\uu, p)$ satisfy the force-free and torque-free
conditions, meaning that the force and torque from the hydrodynamic
stress on each particle balance the total force and torque coming from
the hydrophobic potential and repulsion. We have
\begin{equation}
  \label{eq:force}
 \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i, \quad
 \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu) \, \dif s = T_i,
  \quad i=1,\ldots,N_b,
\end{equation}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor.

To perform a single time step, we
(i) solve~\eqref{eq:SL}--\eqref{eq:SLbc} for $u$,
(ii) compute the force and torque~\eqref{eq:FandTdef},
and finally (iii) solve the Stokes equations~\eqref{eq:stokes}--\eqref{eq:stokesFarfield}
subject to~\eqref{eq:rigid-vel} and \eqref{eq:force}. 
The velocities $(\vv_i, \omega_i)$ obtained
from~\eqref{eq:rigid-vel} are used to update the
particle positions and orientations
using the second-order Adams-Bashforth scheme
with time step size $\Delta t.$

The interactions
generated by the HAP-mobility problem formulation
\eqref{eq:stokes}--\eqref{eq:force} lead to particle self-assembly.
In terms of scaling, $\rho$ sets the effective distance of the hydrophobic
interactions; $\rho_0$ determines the distance where attraction and repulsion are in
balance i.e. smaller values of $\rho_0$ lead to more compact particle assemblies.
The rate of self-assembly are proportional to tension $\gamma$,
inversely proportional to viscosity $\mu$, and
approximately inversely proportional to $\rho$~\cite{Fu20}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral representations}
It is challenging to solve the Stokes and screened Laplace equations
and far-field boundary conditions accurately. 
We recast~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
and~\eqref{eq:SL}--\eqref{eq:SLbc} as boundary integral equations (BIEs)
and discretize each BIE at $N$ points on each of the $N_b$ particles
with a collocation method. To express the solution of
\eqref{eq:SL}--\eqref{eq:SLbc}, we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\xx}) = \DD[\sigma](\xx) = \int_{\partial \Omega} 
  \pderiv{G(\xx-\yy)}{\nnu_\yy}\sigma(\yy)\, \dif s_\yy, 
  \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental
solution to the screened Laplace equation~\eqref{eq:SL}, $K_0$ is the
zeroth-order modified Bessel function of the second kind, $\nnu_\yy$ is
the unit outward normal at $\yy$, and $\sigma$ is a scalar-valued
density function. The subscript in $\dif s_\yy$ denotes integration with
respect to $\yy \in \partial \Omega$. To satisfy the boundary
condition~\eqref{eq:SLbc}, the density function must satisfy\cite{Hsiao2008}
\begin{align}
  \label{eq:SL_BIE2}
  g(\xx) = \frac{1}{2} \sigma(\xx) + \DD[\sigma](\xx), \quad
    \xx \in \Gamma.
\end{align}

For the velocity, we use the representation
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DDD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
where $\eeta$ is a vector-valued density function and
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{(\xx - \yy) \cdot \nnu_\yy}{|\xx - \yy|^2}
    \frac{(\xx - \yy) \otimes (\xx - \yy)}{|\xx - \yy|^2}
    \cdot \eeta(\yy)\, \dif s_\yy.
\end{align}
The Stokeslets and Rotlets are
\begin{align}
  \SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} +
    \frac{\rr \otimes \rr}{|\rr|^2}\right), \quad 
  \RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}, 
\end{align}
respectively, where $\rr = \xx - \aa_i$~\cite{pow-mir1987}. Letting
$\xx$ approach $\Gamma_i$ in~\eqref{eq:velocity}, applying the jump
condition of the double layer potential~\cite{poz1992}, and imposing the
no-slip boundary condition~\eqref{eq:rigid-vel}, the density function
$\eeta$, translational velocity $\vv_i$, and rotational velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \label{eq:vel_BLM_rep}
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DDD[\eeta](\xx) 
    + \sum_{j=1}^{N_b} 
    \left(\SS(\xx,\aa_j) \cdot \FF_j + \RR(\xx,\aa_j) T_j\right),\\
  \label{eq:vel_BLM_rep2}
  \int_{\Gamma_i} \eeta \, \dif s &= \mathbf{F}_i, \quad
  \int_{\Gamma_i} \eeta \cdot (\xx-\aa_i)^\perp \, \dif s = T_i,
\end{alignat}
for $\xx \in \Gamma_i$ and $i = 1,\ldots,N_b$.
%Equation~\eqref{eq:vel_BLM_rep}--\eqref{eq:vel_BLM_rep2} does have a
%unique solution, but we note that
Alternative full-rank layer potential
representations for rigid body motions are possible~\cite{rac-gre2016,
cor-gre-rac-vee2017}.

We discretize~\eqref{eq:SL_BIE2}
and~\eqref{eq:vel_BLM_rep}--\eqref{eq:vel_BLM_rep2} using high-order
interpolation-based quadrature rules. Integrals that are smooth are
computed with the spectrally-accurate trapezoid rule, and
nearly-singular integrals, caused by close contact between two
particles, are computed with a high-order interpolation-based quadrature
rule~\cite{qua-bir2014}.

After discretizing and applying quadrature, the result is an $NN_b
\times NN_b$ and $2NN_b \times 2NN_b$ linear system
for~\eqref{eq:SL_BIE2}
and~\eqref{eq:vel_BLM_rep}--\eqref{eq:vel_BLM_rep2}, respectively.
These are solved with matrix-free GMRES, and we guarantee that the
number of GMRES iterations is mesh-independent by using second-kind
BIEs.

%Once the translational and angular velocities $\vv_i, \omega_i$
%have been determined, a second-order Adams-Bashforth scheme updates the particle
%positions and orientations:
%\begin{equation}
%  \label{eq:adams-bashforth}
%  \begin{aligned}
%    \aa_i(t_{n+2}) &=
%\aa_i(t_{n+1})
%+ \tfrac{3}{2}\Delta t \vv_i(t_{n+1})
%- \tfrac{1}{2}\Delta t \vv_i(t_{n}),\\
%    \theta_i(t_{n+2}) &=
%\theta_i(t_{n+1})
%+ \tfrac{3}{2}\Delta t \omega_i(t_{n+1})
%- \tfrac{1}{2}\Delta t \omega_i(t_{n}).
% \end{aligned}
%\end{equation}

\section{Tunable hydrophobicity}
\label{sec:tunable_hydrophobicity}
The parameters are modeled after those for
phospholipid in water~\cite{Boal}.
We use $\mu = 1$~mPa~s for the viscosity of water
at room temperature. Pure lipid components give a range of interfacial
tensions $0.7$--$5.3$~pN~nm$^{-1}$~\cite{KUZMIN2005, Petelska2012,
  Jackson2016, GarciaSaez}.
we use $\gamma=4.1$~pN~nm$^{-1}$ since
this value gave good agreement with elastic moduli of lipid bilayers~\cite{Fu20,Fu2022_JFM}.
The particle radius $c = 1.25$~nm represents one half 
of phospholipid length~\cite{Boal}, and the screening length $\rho = 5$~nm
derives from experimental force-distance measurements of hydrophobic
attraction~\cite{ErLjCl89,Lietal05,Israelachvili80,Jackson2016}. 
A repulsion modulus $M = 2$~pN and 
distance $\rho_0 = 0.5$~nm gives an interparticle distance around one
particle radius, ensuring the accuracy of the boundary integral
representations~\eqref{eq:SL_BIE} and~\eqref{eq:velocity} without 
overly aggressive mesh refinement. The characteristic time
and length are $1$~ns and  $1$~nm respectively.

\begin{figure}[t!]
\begin{center}
  \includegraphics[width=\textwidth]{Figures/Figure0.pdf}
\end{center}
\begin{caption}{\label{fig:boundary_conditions} The leftmost diagram
  illustrates the particle $\Gamma_i$ with center $\aa_i$, radius $c$,
  and director $\dd$, along with the angle $\alpha$.
  The three rightmost panels plots show $g(\xx)$ and
  the respective solutions $u(\xx)$ of~\eqref{eq:SL} 
  for an isolated particle.}
\end{caption}
\end{figure}
\subsection{Boundary conditions and equilibrium configurations}
The boundary condition $g(\xx)$ in
\eqref{eq:SLbc} defines the spatial distribution of
hydrophobicity and hydrophilicity.
Referring to Figure \ref{fig:boundary_conditions},
\begin{align}
  \label{eq:bc-type}
g(\xx) = a(b + \cos \alpha),\quad a = (\pi c(2b^2 + 1))^{-1/2},\quad \xx \in \Gamma_i,
\end{align}
where $\alpha$ is the angle between the vector $\xx - \aa_i$ and the
particle director $\dd_i = (\cos \theta_i, \sin \theta_i)$
The scalar $a$ normalizes $g$ so that $\int_{\Gamma_i} g^2(\xx) \, \dif s = 1$.
The side of the particle where
$\alpha = 0$ is called the tail and the side where $\alpha = \pi$ is the head.

Based on the choice of $b$,
the boundary condition (BC) \eqref{eq:bc-type}
can be classified into one of three categories:
(Type I, amphiphile, $b = 1$) $g$ is positive on 
one side representing a hydrocarbon-water interface, and
is zero on the other representing an apolar, hydrophilic region;
(Type II, asymmetric hydrophobe, $b = 2$) $g$ is positive
on both sides but greater on one side;
(Type III, poler, $b = -1/2$) $g$ is positive on one
side and negative on the other, representing a
water structure with positive/negative charge~\cite{MaRa76, Ma77}.

Figure~\ref{fig:relax} shows the steady state in a quiescent flow.
Three distinct phases emerge. For Type I, amphiphiles,
the tail interactions are attractive, and particles collectively
form disjoint bilayer components (Figure~\ref{fig:relax}(b)).
We refer to this as the ``bilayer'' phase.

\begin{figure}[t!]
\begin{center}
  \includegraphics[width=\textwidth]{Figures/Figure1and2.pdf}
%  \includegraphics[height=0.3\textheight]{Nb198a_inset.pdf}
%  \includegraphics[height=0.3\textheight]{Nb198b_eta_inset.pdf}\\
%  \includegraphics[height=0.3\textheight]{Nb198c_eta_inset.pdf}
%  \includegraphics[height=0.3\textheight]{Nb198d_eta_inset.pdf}
\end{center}
\begin{caption}{\label{fig:relax}
  Panel (a) shows the initial $N_b = 198$ particle configuration with random
  position and orientation.
  Panels (b), (d), (e) show bilayer, multilamellar, and striated phases
  composed of amphiphilic (Type I), asymmetric hydrophobic (Type II),
  and poler (Type 3) JP, respectively. 
  Panels (c) and (f) plot the free energies~\eqref{eq:free_energy}
  and alignment order parameter~\eqref{eq:S2} with respect
  to $t$. The solid markers in panel (c) are 
  normalized by $60/198$, showing that the energy
  approximately scales with the number of particles.
  Alignment in the striated phase is particle-number dependent; 
  triangle markers, panel (f).}  
\end{caption}
\end{figure}

For Type II, asymmetric hydrophobes,
both sides of the JP are hydrophobic but more so on the tail.
Over short times, these particles self-assemble into bilayers. But unlike for
Type I JP, the heads are also hydrophobic and so over long times, 
the bilayers sort into a ``multilamellar'' phase.
The number of layers depends primarily on the number of particles.
Figure~\ref{fig:relax}(c), for example, shows a multilamellar structure
with four layers and one with two layers in the inset.
Onion-like dendrimersomes have previously been studied by molecular dynamics
simulations using an 
anisotropic pair potential~\cite{C9NR05885K,HongCacciutoLuijtenGranick2008}.

Finally, Type III, bipolar JPs possess a
head that repels the tail of neighboring particles.
These JP initially form chains with
their directors perpendicular to the length of the chain. The chains
form stria where the particles lie on a square grid and the orientations
alternate between layers (Figure~\ref{fig:relax}(d)).
We refer to this as the ``striated'' phase.

%\begin{figure}
%  \begin{center}
%\includegraphics[width=1.0\textwidth]{Figures/Figure2.pdf}
%  \end{center}
%  \vspace{-20pt}
%  \caption{\label{fig:relax_energy}
%  Panel (a) shows that the free energy~\eqref{eq:free_energy} decreases
%  in the mobility problem formulation. The initial energies are for the
%  configuration in Figure~\ref{fig:relax}(a) and the final energies are
%  for the near-equilibrium configurations shown in
%  Figures~\ref{fig:relax}(b)--(d). The solid curves are the energies for
%  the $N_b=60$ cases and the dashed curves are the energies for the $N_b
%  = 198$ cases normalized by $60/198$, showing that the energy
%  approximately scales with the number of particles.  Panel (b) shows
%  the orientational parameters $\tilde{S}_2$ for all 6 relaxation runs.
%  The curve styles are identical those in panel (a). The
%  near-equilibrium configurations depend on the initial configurations
%  and the results show that for both $N_b=60$ and $N_b=198$ cases,
%  bilayer and multi-lamellar configurations have similar trends in
%  orientational parameters. The results of the striated configuration,
%  however, depend on the initial geometrical setup.}
%\end{figure}

\section{Measuring deformation}
To quantify the hydrodynamics of JP phases,
we use the free energy $F$, a strain parameter $E$, and
a scalar order parameters $S_{2}$ for alignment.
First we simplify the form of the free energy~\eqref{eq:free_energy}.
Using integration by parts and~\eqref{eq:SL}, we obtain
\begin{equation}
\label{eq:free_energy2}
F = -\gamma
\int_{\partial \Omega} \rho g \nabla u \cdot \nnu \,\dif s
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right).
\end{equation}
%
Here, we have substituted $g$ for $u$ since the boundary values are
given. However, evaluating $\nabla u \cdot \nnu$ on $\partial \Omega$ based
on~\eqref{eq:SL_BIE} involves calculating a quadruple layer potential
which has a well-known obstacle in numerical implementation.
Appendix \S\ref{sec:appendix} describes
how we overcome this obstacle.

Figure~\ref{fig:relax}(c) tracks the free energy profiles for all
relaxation runs.  In quiescent background flow, the energies are
decreasing with respect to $t$ showing time stepping correctly
accounts for viscous dissipation.  Furthermore, the normalized energy
plots provide evidence that the free energy per particle with specified
boundary condition is independent of the total particle number $(N_b)$.

To measure positional order, we introduce the strain
parameter, 
\begin{equation}
\label{eq:SP}
E = \frac{1}{N_b} \sum_{i=1}^{N_b}
\left\|\tfrac{1}{2}(\mathsf{F}_i^T \mathsf{F}_i - I)\right\|,
\end{equation}
where $\mathsf{F}_i$ is an approximate deformation gradient
and $\| \cdot \|$ is the Frobenius norm. 
The Green-Lagrange strain tensor
$\tfrac{1}{2}(\mathsf{F}_i^T \mathsf{F}_i - I)$
measures departure of fluid deformations from a rigid
body motion.  To define $\mathsf{F}_i$,
we
solve the overdetermined system 
\begin{align}
\aa_j(t) - \aa_i(t) = \mathsf{F}_i(\aa_j(0) - \aa_i(0)),\quad j =
  1,\ldots, N_b,
\end{align}
for $\mathsf{F}_i$ by weighted least squares.
That way, if the particle positions are given by
a map $\ff(\aa_i(0),t) = \aa_i(t)$, then
$\mathsf{F}_i \approx \nabla \ff(\aa_i(0),t)$.
The weights $w_i =
\exp(-\|\aa_j(0) - \aa_i(0)\|/4c)$ with particle radius $c$ ensure that
the linear approximation holds for particles near $\aa_i$.

Finally, we use the scalar order parameter $S_2$
to quantify the orientational order\cite{Selinger2016}:
\begin{equation}
  \label{eq:S2}
S_2 = \frac{1}{N_b} \sum_{i=1}^{N_b} \frac{1}{2}(3\cos^2(\theta_i - \bar \theta) - 1).
\end{equation}
Here $\bar \theta_i$ is a circular mean defined as the orientation of the
principal eigenvector of the matrix
$\sum_{j} \dd_j\dd_j^\top$ where $j$ runs over the particles
whose centers lie with $4c$ of that of particle $i$. 
Defined as such, $S_2$ lies in the range $-1/2 \le S_2 \le 1$.
with a value $S_2 = 1$ indicates perfect alignment between particles
while $S_2=-1/2$ indicates isotropic alignment locally.
This cutoff distance was chosen empirically
so that the average includes nearest neighbors
but excludes next nearest neighbors.

In Figure~\ref{fig:relax}(f), the multilamellar phase is highly ordered 
whereas the bilayer phase is somewhat
disordered because it consists of several components forming isolated
bilayers, micelles, and vesicles (Figure~\ref{fig:relax}(b)).

The striated phase order admits more than one pattern
depending on the number of particles (Figure~\ref{fig:relax}(f), bottom and top curves).
When the number of particles is fewer, there is only a
single pattern in particle orientations where the directors are more or
less parallel and alternate directions (Figure~\ref{fig:relax}(e),
inset). Larger number of particles results in multiple alignment patterns:
the alternating sign pattern as in the small particle number case
(Figure~\ref{fig:relax}(e), top right rectangle) and an 'X'-shaped
alignment (Figure~\ref{fig:relax}(e), top left rectangle).

%%%%%%%

\section{Results and Discussion}
\label{sec:results}
\begin{figure}[t]
\begin{center}
\includegraphics[width=\textwidth]{S2ESummary.pdf}
\end{center}
\caption{\label{fig:s2esummary}asdf}
\end{figure}
We subjected the Type I, II, and III JP phases to background flows and measured 
their material response.  The background flows were a shear flow
\begin{equation}
\label{eq:sh_flow}
\uu_\infty^{\text{sh}}(\xx) = 
\dot\gamma (\ee_y \cdot \mathbf{x})\ee_x,
\end{equation}
and a Taylor-Green (TG) flow
\begin{equation}
\label{eq:tg_flow}
\uu_\infty^{\text{TG}}(\xx) = 
\lambda \dot \gamma \left(
-\cos(\ee_x\cdot\xx/\lambda)\sin(\ee_y\cdot\xx/\lambda)\ee_x
+\sin(\ee_x\cdot\xx/\lambda)\cos(\ee_y\cdot\xx/\lambda)\ee_y\right),
\end{equation}
where the orthonormal vectors $\ee_x$ and $\ee_y$ are the horizontal
and vertical directions, respectively. 
In both cases, $\dot \gamma$ is the shear rate.
To obtain a range for $\dot \gamma$, 
consider the capillary number 
\begin{equation}
\label{eq:capillary}
C_a = \frac{\mu \dot \gamma R^3}{\kappa}
\end{equation}
where $R = 10$ nm and $\kappa = 20\;\kbt$ are 
the characteristic phase sample radius 
and bending rigidity, respectively. 
That is, $\kappa$ gives the elastic stress 
for restoring the phase-sample shape.
Shear stress overcomes elastic stress   
when $C_a > 1$ and this yields 
$\dot \gamma > 2 \cdot 10^{-2}$ ns$^{-1} = 2 \cdot 10^6$ s.
Thus our values for $\dot \gamma$ range between 
$0$ ns$^{-1}$ corresponding to quiescent flow
up to $0.2$ ns$^{-1}$ capable of rupturing the sample. 
Such shear rates are compatible with molecular dynamics 
simulation \cite{Bradner}, those for large unilamellar vesicles (LUV),
\cite{Bernard2005ShearinducedPA} and shearing in blood capillary vessels 
 where shear rates range over 100--1000 s$^{-1}$ \cite{NEEDED}.

The shear flow replicates the motion of a fluid between two parallel, moving plates
excluding wall effects.  In TG flow, the flow pattern
consists of an array of counter rotating cells each with size 
$2\pi \lambda \times 2\pi \lambda$.  Throughout, we use $\lambda = 2$ nm
so that the JP phases occupy about nine cells.  To compare shear
on equal terms, the flow rate $\lambda \dot \gamma$  
was chosen for \eqref{eq:tg_flow} so that it has has the same
rate viscous dissipation per area as for \eqref{eq:sh_flow}.
In other words, 
$\int_{A} \frac{1}{2}\mu |\nabla \uu + \nabla \uu^T|^2 \,\dif A$ 
is the same for 
$\uu = \uu_{\infty}^{\text{sh}}$ as for 
$\uu = \uu_{\infty}^{\text{TG}}$
when integrated over a common cell $A$.

For initial data, the steady-state Type I amphiphiles phase
consisted of a number of disjoint bilayer components
(Figure \ref{fig:relax}(b)). To also have a single-component Type I phase, 
we reinitialized the data to have a ring-shaped vesicle bilayer.
This phase is distinguished from "bilayer" by "vesicle BL" in the figures.
For the same number of JP, the vesicle bilayer
has lower free  steady-state energy ($F = 24.6 \;\kbt$ nm$^{-1}$) than the 
disordered bilayer phase ($F = 24.6 \;\kbt$ nm$^{-1}$). 
Since the simulations are for two dimensions,
multiplying $F$ by a length gives the energy 
of a transversely invariant, three-dimensional 
phase.

Figure \ref{fig:s2esummary} (e)--(l) show the typical streamlines
for JP phases in background flow.  Although the plots only show
the region near the JP phases, the fluid velocity $\uu$ and 
order parameter $u$ are solved for in all of $\mathbb{R}^2$.
Moreover, $\uu$ satisfies the rigid boundary condition at 
every JP-fluid interface i.e., the particles interact with local
flow field and are not merely carried by background flow.

The simulation setup consisted of placing each of the three JP phases
in either a shear or TG background flow and then solving for the 
particle trajectories in time.  A total of $500$ steps were solved
for with $\Delta t = 2$ ns, yielding time courses for various 
$\dot \gamma$ up to $T = 1$ $\upmu$s.  We then post-processed the trajectories
to infer free energy $F$, alignment order $S_2$, and strain $E$.

\subsection{Strain and alignment}
The plots reveal highly distinct material responses from each phase.
Figure \ref{fig:s2esummary} (a) and (b) show the time-averaged
strain $\frac{1}{T}\int_0^T E \,\dif t$ over a broad
range of shear rates.  Strain $E$
measures the non-rigid-body deformation relative to the initial state
and expectedly increases with shear rate.
The greatest increase occurs with the Type I amphiphiles, 
suggesting that this phase has the lowest elastic modulus. In comparison, 
the  striated phase consisting of  Type III polar JP is
basically rigid.

Figure \ref{fig:s2esummary} (c) and (d) plot the time-averaged
alignment order $S_2$. The vesicle and striated phases possess
the greatest alignment order and, 
for the most part, $S_2$ decreases with increases in shear rate.
But there is an exception.  The Type I bilayer phase alignment
order actually increases under both shear and TG flows (solid squares).  
This suggests that 
the mixing action of background flow has the effect of moving the
disordered bilayer phase out of a local equilibrium and into a state
with greater order.
This surprising increase, 
brought about by combining previously broken bilayer 
end caps, is accompanied
by a slight decrease in free energy (Figure \ref{fig:fsummary}a, solid squares).

The sudden jumps in  Figure \ref{fig:s2esummary} (a) and (b)
and commensurate drops in panels (c) and (d) indicate
the existence of a critical shear rate $\gamma_*$.  This means that 
for $\gamma < \gamma_*$, the JP phase remains intact while
for $\gamma > \gamma_*$, the phase ruptures. 
For example, the vesicles phase has 
$\gamma_* = 0.05$ ns$^{-1}$ 
under shear flow and TG flow. However, the critical shear rate
is flow-pattern dependent because 
the striated phase has $\gamma_* = 0.14$ ns$^{-1}$ under shear flow
while $\gamma_* = 0.05$ ns$^{-1}$ under  TG flow.
Due to variable stress distribution, 
the critical shear rate in TG likely 
depends on the cell size parameter $\lambda$.
The disordered bilayer phase does not really possess a critical
shear rate because it already consists of several, disjoint components.
The jump in Figure \ref{fig:s2esummary} (a) (solid squares)
simply corresponds to disjoint components drifting apart in the shear
flow.  The Type II multilamellar phase did
not rupture for any of the shear rates considered.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\textwidth]{FSummary.pdf}
\end{center}
\caption{\label{fig:fsummary}}
\end{figure}

\subsection{Free energy}
Figure \ref{fig:fsummary} plots the time-averaged
relative free energy $\frac{1}{T}\int_0^T F - F_0 \,\dif t$ where $F_0$ is 
\eqref{eq:free_energy2} at $t = 0$.  The 
greatest increase occurs with the Type II 
multilamellar phase, with the other phases 
showing a moderate change in a few \KBT\;per 
nm.

Figure \ref{fig:fsummary} (c) plots free energy 
against $x = c E$ representing the mean  
inter-particle displacement;    
zero displacement corresponds
to an inter-particle distance where 
repulsion and attraction are in balance;
$c = 1.25$ nm is particle radius.
The energies in Figure \ref{fig:fsummary} (c) are essentially  
quadratic in $x$ suggesting that the interactions 
are harmonic locally around equilibrium.
The harmonic bond strength is strongest between 
polar JP, and greater by a factor of two and five than
that for asymmetric hydrophobes and amphiphiles,
respectively.  This makes sense when considering 
that the free energy of the steady-state striated
phase is also greatest (Figure \ref{fig:relax} (c), around 85 \KBT\;nm$^{-1}$).
However, the bond strength in the multilamellar 
phase comes in second, despite having least steady-state free energy
(around 15 \KBT\;nm$^{-1}$).
The data point to the fact that the binding properties 
do not merely scale with free energy, 
but also involve the details of the interface's
hydrophobicity 
which are tuned by \ref{eq:bc-type} in our model.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\textwidth]{RheologySummary.pdf}
\end{center}
\caption{\label{fig:rheologysummary}}
\end{figure}
\subsection{Rheology}
So far, we have considered static material properties, namely
constant strains maintained by an external force provided by   
background flow. Now we consider dynamic material properties.
Specifically, we look at the strain rate 
\begin{equation}
\label{eq:strainrate}
\dot E = \left(\frac{1}{T} \int_0^T \left|\frac{dE}{dt}\right|^2 \, \dif t \right)^{1/2}.
\end{equation}
The reason for defining $\dot E$ in this was is that 
the shear flow $\uu_{\infty}^{\text{sh}}$
simulations correspond to a complex fluid within a drag plate
rheometer where the shear stress is set by $\mu \dot\gamma$  
and the observed, sample shear rate is $\dot E$. 
The definitions carry over to the TG flow case directly. 
Also, as a practical consideration,
the rotation of non-isotropic JP phases in the background 
flow leads to oscillation in the strain time courses 
making systematic fitting challenging (see 
Supplementary Material Figures \ref{fig:veshraw}--\ref{fig:sttgraw}). 
The integral in \eqref{eq:strainrate} averages out these oscillations
and gives physically meaningful values.

The shear stress--strain rate data are 
excellently fit by power laws.  
Figures \ref{fig:rheologysummary} (a), (b), and (c)
show the data for Type I, Type II and Type III JPs, respectively,
under both shear flow (squares) and TG flow (triangles).
The bilayer phases consisting of Type I JPs are basically Newtonian
with viscosity $0.014$ MPa\;ns $= 14$ mPa\;s,
about 14 times the the viscosity of water at room temperature
(Figures \ref{fig:rheologysummary}(a), dashed curve). 
The vesicle bilayer phase is slightly shear thickening
(Figures \ref{fig:rheologysummary}(a), solid curve).
The multilamellar phase consisting of Type II JPs
is somewhat shear thinning,  but its viscosity 
is larger than for Type I.  Finally, the striated
phase is strongly shear-thinning.  This 
further explains the rigid-like behavior of the 
striated phase found in the strain, shear-rate relationships from 
Figure \ref{fig:s2esummary}(a), (b) (triangles).

\begin{figure}[t]
\begin{center}
  \includegraphics[width=\textwidth]{critical_shear_and_cell_size.pdf}
\end{center}
\caption{\label{fig:PramChange}
In panel (a), relative change in $\dot \gamma_*$ is linear 
in the relative change of tension with slope $1$.
Conversely, relative change in $\dot \gamma_*$ 
has slope $-2$ with respect to relative change in tension.
Panels (b)--(e) show the tank-treading configuration 
of a vesicle bilayer the configurations with streamlines of a vesicle is in a TG flow at
$t=1$ $\upmu$s with $l= 1,2,3,4$, respectively. In panel (e), the
  relative enclosed area $A/A_0$ calculated from the bilayer midplane
  decreases over time. The inset shows the relative length $L/L_0$ which
  remains nearly constant with small magnitude of oscillations. The
  inset and main panel have the same horizontal axis.}
\end{figure}

\subsection{Discussion}
We chose the range of shear rates $\dot \gamma \in [0, 0.02]$ ns$^{-1}$
by analyzing the capillary number of the JP phases.  Within this range, 
there was a critical shear rate $\dot \gamma_*$ for the vesicle bilayer
and striated phases where shear stresses exceeded elastic stresses,
leading to rupture.  (The multilamellar phase did not rupture.)
By varying model parameters, we can perform a dimensional analysis 
on $\dot \gamma_*$.  Focusing on Type III JP in shear flow, 
$\dot \gamma_* = 0.14$ ns$^{-1}$ for the basic model parameters
(see Section \ref{sec:tunable_hydrophobicity}).  Increasing 
the tension parameters $\gamma$ and $M$ by 5\% (collectively denoted
$\Delta \gamma/\gamma$) led to a 4.29\% increase in critical shear
rate (denoted $\Delta \dot \gamma_*/\dot \gamma_*$).  This suggests
that critical shear rate scales linearly with the tension parameters.
Conversely, increasing the length parameters $c$, $\rho$, and $\rho_0$
by 5 \% (collectively denoted $\Delta c/c$) led to a -11.43 \% 
decrease in critical shear rate.  This points to an inverse-quadratic
scaling in length. Figure \ref{fig:PramChange}(a) summarizes the 
findings and suggest the scaling
\begin{equation}
\label{eq:CSR}
\mu \dot\gamma_* = K_{\text{III}} \frac{\gamma L}{c^2}.
\end{equation}
The scalar $K_{\text{III}}$ is specific to the striated phase under shear flow.
Here $L$ is the diameter of the phase sample, which scales proportionally 
with $\gamma^*$ since the striated phase tends to cleave during
rupture (see Supplementary Movie S4).
The scaling for vesicle bilayers in shear flow is different
and takes the form $\mu \dot\gamma_* = K_{\text{I}} \gamma c^2/(\rho L^2)$
and is proportional to particle size \cite{Fu2022_JFM}, versus inverse quadratic
like for Type III JPs.

The Results section showed that 
variations in free energy and strain 
generally depend the distribution
of fluid stresses i.e. shear versus TG flow.   
To further illustrate how fluid stresses interact with the geometry of the JP
phases, we considered the initially circular vesicle bilayer
for various TG-flow cell sizes (Figure \ref{fig:PramChange}(b)--(e)). 
In all cases, the vesicle and TG flow reach a steady-state tank-treading motion,
but the geometry and motion varies nonmonotonically. 
For $\lambda$ ranging from $1$ nm to $4$ nm,
the steady-state vesicle shape goes from an octagon, to a dimpled square,
to a rounded diamond. \hl{Surprisingly, the direction of rotation switches,
so that for $\lambda = 1$ nm, the rotation is counterclockwise while
for $\lambda = 3$ nm, the rotation is clockwise, for example.  }
Finally, we calculated the vesicle area $A$ and length $L$ (initial area $A_0$,
initial length $L_0$). 
As shown in Figure~\ref{fig:PramChange}(f), the relative
area $A/A_0$ achieves a steady-state that is greatest for $\lambda = 4$ nm
yet least for $\lambda = 2$ nm. 
(The relative length $L/L_0$, however, is
constant in $t$ for all cell sizes (Figure~\ref{fig:PramChange}(e),
inset)).

We have excluded thermal fluctuations because they are unresolved in the JP phases we consider. 
For a tensionless, planar membrane, the spectrum of the Fourier coefficients for the 
profile height 
$h(\xx) = \sum_{\kk} h(\kk)e^{i \kk \cdot \xx}$ satisfy\cite{Merath2006FluctuationSO}  
\begin{equation}
\label{eq:fourier_spectrum}
\langle |h(\kk)|^2 \rangle = \frac{\kbt}{L_xL_y}\frac{1}{\kappa k^4}
\end{equation}
where $\xx$ is a point in an $L_x$ by $L_y$ rectangular membrane patch,
$\kk = \left(2\pi m/L_x, 2\pi n/L_y\right)$ with
$m, n = 0, \pm 1, \pm 2, \dots$ not both zero, and $\kappa$ is the membrane bending modulus.

To estimate $h(\kk)$, 
the $N_b = 60$ JP
vesicle in Figure \ref{fig:ES2_summary}(e) 
has arc length $L_x \approx 50$ nm. 
Our prior work\cite{Fu20,Fu2022_JFM} gave $\kappa  \approx 20$ $\kbt$.
Since our simulations are two-dimensional, we restrict our attention
to transversely invariant fluctuations where $n = 0$.
Substituting these values gives 
\begin{equation}
\label{eq:fourier_spectrum_upper}
\langle |h(\kk)|^2 \rangle \approx \frac{4 \text{ nm}^3}{L_y}\frac{1}{m^4}, \quad m = \pm 1, \pm 2, \dots
\end{equation}
For any moderate width, say $L_y > 4$ nm, the leading Fourier coefficient
has an expected magnitude less than $1$ nm with the remaining coefficients
dropping off to zero precipitously.  The expected amplitudes coming from thermal
fluctuation are therefore negligible compared to particle size.
The elastic
moduli for the Type (II) and Type (III) phases are by greater by a factor of two
and five, respectively, than for Type (I) (Figure \ref{fig:F_summary}(c)), making thermal fluctuation
even more negligible in these cases.

The fact that the fluctuation amplitudes are small 
is not a consequence of stiffness. Quite the opposite, the value $\kappa = 20$ $\kbt$
is in  agreement with the experimental literature where 
micron sized vesicles observably fluctuate. 
Rather, larger energies are required to fluctuate membranes 
over smaller wavelengths and these energies exceed available 
thermal energy when the wavelengths approach tens of nanometers
as demonstrated above.

The calculated viscosities---14 mPa s for Type I 
and 21 mPa s for Type II JP phases---are physically reasonable.
In two-dimensions, the Einstein viscosity correction 
takes the form $1 + 2.0 \phi + 4.0\phi^2 + O(\phi^3)$
in the volume fraction $\phi$
\cite{PhysRevFluids.1.043301,Haines2011APO,Khair2006TheC,Brady1983TheEV}.
For the JPs in Figure \ref{fig:s2esummary}(e)--(h), this volume fractions
is about $0.5$ and would give a small viscosity $3$ mPa s.
On the other hand, the Einstein correction is only valid for dilute
suspensions, and for large volume fractions like $\phi = 0.5$, 
one expects viscosities ten or more times that of water  
\cite{Konijn2014ExperimentalSO}.

Nevertheless, further simulations results are required to 
differentiate between purely suspension effects
and particles' preferred alignments
in analyzing viscosities from Figure \ref{fig:rheologysummary}. 
For example, the difference between the values 14 mPa s for Type I and 
21 mPa s for Type II JPs can be explained by the fact that 
the multilamellar phase is more tightly packed than the bilayer phase
(Figure \ref{fig:s2esummary}(e)--(h)).
Furthermore, it seems that the viscosity, and the tendency toward
shear thinning in the case of Type III JPs where viscosity is 
effectively infinite for low strain rates 
(Figure \ref{fig:rheologysummary}(c)), is related 
to particle coordination.  
To leading order, a single particle coordinates with: 
one neighbor in the apposing bilayer for Type I, two neighbors
at the tail and head interfaces for Type II, and four
neighbors (front, back, left, and right) for Type III.

\section{Conclusion}
\label{sec:conclusion}

In this work, we employ the newly developed JP model using
BIEs~\cite{Fu20, Fu2022_JFM} and tune the boundary conditions with
energy normalization to study the collective dynamics of amphiphilic
(BC(i)), biased hydrophobic (BC (ii)), and bipolar (BC(iii)) JPs under
various flowing conditions (quiescent flow, linear shear flow, and
Taylor-Green flow). Three quantities are computed to characterize the
dynamics of the collective configurations of JPs: free energy $F$,
strain parameter $E$, and scalar order parameter $S_2$. 
%We quantify how
%the collective hydrodynamics of JPs is affected by an external flow.
Under a given flow, we use these three measures of deformations to
quantify the differences in the collective dynamics between the three
types of JPs. These results, summarized below, provide general insight
into the dynamic control of active particles in a viscous suspension.

Within the framework of boundary integral equations, we derive a
convenient expression for computing the free energy using the single
layer potential (see equation~\eqref{eq:normal_deriv}). In a quiescent
flow, the free energy profiles demonstrate that the relaxation process
for particles confined in a certain size of box is independent of the
number of particles. However, we find that the final configurations does
depend on the initial distribution of particle directors $\dd_i$.
Therefore, multiple patterns or local energy minimum states may appear
depending on the initial setup.

Under a relatively weak linear shear flow, the amphiphilic JPs behave as
a unilamellar vesicle that elongates and tank-treads, with the scalar
order parameter increasing over time. The assembly of multi-lamella and
striated JP structures resemble a rigid body motion with minimal
deformation. The effective viscosity of the material quantitatively
validates this result. High shear-rate cases provide a range of critical
shear rates where the structures break apart and undergo topological
changes. We further show that free energy, scalar order parameter, and
strain are effective measures to quantitatively capture the collective hydrodynamics of
JPs under a linear shear flow.

Under a Taylor-Green flow, the amphiphilic JPs are the most interesting,
exhibiting different vesicle shapes depending on the ratio of the size
of the TG flow to the vesicle size. These results show that the shape of
the vesicle, whether square or polygonal, can be controlled by adjusting
the size of the cell in the TG flow. On the other hand, the assembly of
multi-lamella JPs and the striated JPs behave more like a rigid body
with connected subdomains, and the number of subdomains increases with
increasing strength of the TG flow. Overall, the multilamellar (BC (ii))
JP assembly behaves as a shear-thinning fluid, while the striated (BC
(iii)) JP assembly  possesses a yield stress.

The results reported here provide an inroad modeling framework for
hydrodynamics of active colloids~\cite{Meredithetal2022,
McGlassonBradley2021, Vutukuri2020, Mallory2017}. The present study also
helps us understand the rheology of JP oligomers that may be realized in
the experiments. We are extending this study to three dimensional
systems with more realistic features such as size distributions of JPs
and thermal fluctuations~\cite{kohl-cor-che-vee22}. From a numerical
perspective, it is straightforward to include random perturbations in
the particle shape and boundary condition that mimic interfacial
properties found in lab conditions~\cite{Bradley2016, Bradley2017,
Zarzaretal2015, doi:10.1021/la503455h}.

\section{Appendix}
\label{sec:appendix}
The boundary integral calculation
of \ref{eq:free_energy2} relies
on the following identiy 
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu(\xx)=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \partial \Omega.
\end{equation}
%
Here, $\tt(\xx)$ is the tangent vector and $\dif/\dif s$ is the
arclength derivative.
Substituting~\eqref{eq:normal_deriv}
into~\eqref{eq:free_energy2} for the normal derivative leads to the
single layer, $\SS$, which is more straightforward to evaluate.
The arclength derivative is computed using
the spectrally accurate Fourier transform.

To prove ~\eqref{eq:normal_deriv}, let
\begin{equation}
  \label{eq:SLP}
  \mathcal{S}[\sigma](\xx) = \int_\Gamma G(\xx-\yy) \sigma(\yy)\, \dif s_\yy
\end{equation}
be the single layer potential for a density function $\sigma.$ Fix $\xx
\in \Gamma$, let $\nnu_{\xx} = \nnu(\xx)$ and $\nnu_{\yy}$ be the unit
normal at $\xx$, respectively $\yy$, in $\Gamma$, and let $\zz \in
\Omega$. The subscripts in $\nabla_{\zz}$ and $\nabla_{\yy}$ denote
differentiation with respect to $\zz$, respectively $\yy$.

Recall from~\eqref{eq:SL_BIE} that $u = \mathcal{D}[\sigma]$.
Then  
\begin{align*}
\nabla_{\zz} u(\zz) \cdot \nnu_{\xx}
&=\nnu_\xx \cdot \nabla_\zz \int_\Gamma \frac{\partial G(\zz-\yy)}{\partial \nnu_\yy}\sigma(\yy) \,\dif s_\yy\\
&=\int_\Gamma \nnu_\xx^\top \left(\nabla_\zz\nabla_\yy^\top  G(\zz-\yy)\right) \nnu_\yy\sigma(\yy)  \,\dif s_\yy\\
  &=-\int_\Gamma \nnu_\xx^\top \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\nnu_\yy\sigma(\yy)\ \dif s_\yy,
\end{align*}
since we can interchange $\nabla_\zz$ with $-\nabla_\yy$.
Following~\citet{Hsiao2008}, \S 1.2,
\begin{align}
\nnu_\xx^\top \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\nnu_\yy
=
-\tt_{\xx}^\top \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\tt_{\yy}
+ \Delta_{\yy}G(\zz-\yy) \tt_{\xx}\cdot \tt_{\yy}.
\end{align}
Then, using that $\Delta_{\yy} G(\zz-\yy) = \rho^{-2} G(\zz-\yy)$,
interchanging $\nabla_\yy$ with $-\nabla_\zz$ once more, and integrating
by parts in arclength $s$, we obtain 
\begin{align*}
\nabla_{\zz} u(\zz) \cdot \nnu_{\xx}
&=-\int_\Gamma  \Delta_\yy G(\zz-\yy) {\bf t}_\xx \cdot   {\bf t}_\yy \sigma(\yy)\ \dif s_\yy
+\int_\Gamma({\bf t}_\xx\cdot\nabla_\yy)({\bf t}_\yy\cdot\nabla_\yy G(\zz-\yy))\sigma(\yy)\ \dif s_\yy\\
&= -\int_\Gamma \frac{1}{\rho^{2}} G(\zz-\yy) {\bf t}_\xx \cdot {\bf t}_\yy \sigma(\yy)\ \dif s_\yy  
-({\bf t}_\xx\cdot\nabla_\zz)\int_\Gamma \frac{\dif}{\dif s_\yy}G(\zz-\yy) \sigma(\yy)\ \dif s_\yy\\
&= -\frac{1}{\rho^2} {\bf t}_\xx\cdot \int_\Gamma G(\zz-\yy){\bf t}_\yy \sigma(\yy)\ \dif s_\yy + 
({\bf t}_\xx \cdot \nabla_\zz)\int_\Gamma G(\zz-\yy)\frac{\dif }{\dif s} \sigma(\yy)  \dif s_\yy.
\end{align*}
%
Letting $\zz\to\xx\in\Gamma$, and noting that both sides of the equation
are continuous, we obtain~\eqref{eq:normal_deriv}.

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
%We thank Dr.~S\'ebastien Michelin and Dr.~Cecile Cottin-Bizzonne from LadHyX - Ecole Polytechnique Institut Polytechnique de Paris for organizing the focus-session ``interfacial active matter'' in the conference APS-DFD 2021 and the invitation of submitting a special collection paper. 
We thank useful conversations with E. Corona, M. Rachh, S. Jiang, and Thomas Beck.
B.Q.~acknowledges support from the Simons Foundation, Mathematics and Physical Sciences-Collaboration Grants for Mathematicians, Award No. 527139.
Y.-N.Y.~acknowledges support from NSF (Grants No. DMS 1614863 and No. DMS 195160) and Flatiron Institute, part of Simons Foundation.

% put your acknowledgments here.
\end{acknowledgments}

% Create the reference section using BibTeX:
\bibliography{reference}

\input{Supplementary}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{DRAFT RESULTS BELOW}
{\bf More results:}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.3\textwidth]{CS_E.pdf}
   \includegraphics[width=0.3\textwidth]{CS_LOP.pdf}
    \includegraphics[width=0.3\textwidth]{CS_MSD.pdf}
  \end{center}
\caption{A checkerboard in shear flow with shear rates
  $\dot\gamma=\{0.125, 0.15\}$. We propose the order parameter defined
  by $S_{loc} = \frac12(3\cos(\theta-\bar\theta)^2-1)$}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.3\textwidth]{CTG_E.pdf}
   \includegraphics[width=0.3\textwidth]{CTG_LOP.pdf}
    \includegraphics[width=0.3\textwidth]{CTG_MSD.pdf}
  \end{center}
\caption{A checkerboard in Taylor-Green flows when the flow strength $V_0=\{0.2,0.3\}$. }
\end{figure}

\subsubsection{}

%%%

\section{Governing Equations}
We consider a suspension of $N_b$ Janus particles in the fluid domain
$\Omega$.
The boundary of the $\Omega$ is $\bd\Omega = \Gamma_1 \cup
\cdots \cup \Gamma_{N_b}$, where $\Gamma_i$ is the boundary of Janus
particle $i$. We also consider a bounded fluid domain where $\Gamma_0$
is the fixed boundary of $\Omega$. In this case, $\bd\Omega = \Gamma_0
\cup \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$.

%%%
\subsection{Hydrophobic Attraction Potential Mobility Problem} 
The mathematical formulation is a nonlinear system for the dynamics of a
collection of rigid Janus particles. For the sake of completeness, we
summarize the governing equations here, and a complete description is in
our previous work~\cite{Fu2022_JFM}. The interactions satisfy a
system of partial differential equations (PDEs) that describe the
hydrodynamic interactions and hydrophobic interactions. Assuming
inertial terms are negligible, the solvent is governed by the Stokes
equations
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0},     && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity, $p$ is the pressure, $\uu_\infty$ is the
background flow, and $\mu$ is the (constant) solvent viscosity. The domain $\Omega
= \Omega(t)$ is the fluid region surrounding the particles and changes
shape as the particles move. Since the particles are rigid, the solvent
velocity satisfies the no-slip rigid body motion boundary condition
\begin{align}
  \vv_i + \omega_i (\xx - \aa_i)^\perp = \uu(\xx), 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity and $\omega_i$ is the
angular velocity. Once these velocities are determined, we use
second-order Adams-Bashforth to update the particle configuration.

The translational and angular velocities guarantee that the fluid forces
balance the imposed forces. The imposed forces include the hydrophobic
interactions that arise so that particles minimize the free energy of
the structure of the surrounding water molecules. The free energy
functional is
\begin{align}
\label{eq:free_energy}
  F[u] = C \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} f(u)\right)
  \,d\xx,
\end{align}
where $u$ is an order parameter for the structure of water, $\rho$ is a
decay length, $C$ is a constant, and $f(u)$ is a potential.
%$c$ is ``effective" inter-particle distance below which the steric repulsion becomes important.
Hydrogen-bond persistence times are on the order of picoseconds which is
much smaller than characteristic time for particle motion.
%\cite{MaGa13}. 
Thus we assume $u$ minimizes $F[u]$ for all times. Assuming appropriate
conditions on $f(u)$,
% (\cite{evans10}, \S 8.2),
$u$ satisfies the Euler-Lagrange equation
\begin{alignat}{3}
  \label{eq:SL}
  -\rho^2 \Delta u + \tfrac{1}{2}f'(u) &= 0, && \xx \in \Omega, \\
  u &= g, && \xx \in \bd\Omega, \\
  u &\rightarrow 0, \qquad&& |\xx| \rightarrow \infty.
\end{alignat}
Equation~\eqref{eq:SL} is solved with a boundary integral equation
method. The boundary condition $g$ is a material label that is
transported with the particle.

The particles lower the free energy $F[u]$ of the surrounding water
by moving. We calculate the rate of change of $F[u]$ using
variation of the domain. % ~\cite{Fu2018_SIAM,Bandle2015, Schiffer1954, Grinfeld2010}.
Carrying out this variation yields the stress  
\begin{align}
  \label{eq:stress}
\mathbf{T}
= C \left[ \rho^{-1} f(u) \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 
  2\nabla u \nabla u^T\right)\right].
\end{align}
The imposed forces and torques come from the integration of $\mathbf{T}$
along the particle boundary. These forces and torques couple the Stokes
equations \eqref{eq:stokes}, to semilinear elliptic equation
\eqref{eq:SL}. Solving for the translational and rotational velocities
gives the particle evolution.

To solve the Stokes equation with the correct boundary conditions, we
represent the velocity as the of a layer potential with Stokeslets and
rotlets
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
%
Then, the density function $\eeta$, translation velocity $\vv_i$ and
angular velocity $\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eq:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \cdot \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eq:mobility1}
  \int_{\Gamma_i} \eeta \, \dif s &= \mathbf{F}_i, 
  &&i = 1,\ldots,N_b, \\
  \label{eq:mobility2}
  \int_{\Gamma_i} \eeta \cdot (\xx-\aa_i)^\perp \, \dif s &= T_i,
  &&i = 1,\ldots,N_b.
\end{alignat}

\subsection{Normal Derivative of Double Layer Potential and HAP Energy}

We adopt the double layer potential to express the solution of \eqref{eq:SL} where the condition is $f(u)=u^2$. Then it is given by 
\begin{align}
\label{eq:energy_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy,
\end{align}
%
where $G(\xx)$ is the fundamental solution to the screened Laplace
equation \eqref{eq:SL} and $\nnu_\yy$ is the outward normal.
Since the free energy functional can be written as 
\begin{align}
\label{eq:free_energy2}
  E[u(\xx)] &= C\rho \int_\Gamma u \frac{\partial u}{\partial \nnu}  \,\dif s, \quad \xx\in\Gamma_i,
\end{align}
%
we can then refer to the identity (\cite{Hsiao2008}, \S 1.2) to calculate this boundary integral equation without evaluating the normal derivative of the double layer potential directly.

Consider
\begin{align}
\nnu_\xx \cdot \nabla_\zz u(\zz) &=\nnu_\xx \cdot \nabla_\zz \int_\Gamma \frac{\partial G(\zz-\yy)}{\partial \nnu_\yy}\sigma(\yy) \,\dif s_\yy\\
&=\int_\Gamma \nnu_\xx\cdot \left(\nabla_\zz\nabla_\yy^\top  G(\zz-\yy)\right)\cdot \nnu_\yy\sigma(\yy)  \,\dif s_\yy\\
&=-\int_\Gamma \nnu_\xx\cdot \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\cdot\nnu_\yy\sigma(\yy)\ \dif s_\yy\\
&=-\int_\Gamma {\bf t}_\xx\cdot\Delta_\yy G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy+\int_\Gamma({\bf t}_\xx\cdot\nabla_\yy)({\bf t}_\yy\cdot\nabla_\yy G(\zz-\yy))\sigma(\yy)\ \dif s_\yy\\
&= -\int_\Gamma {\bf t}_\xx\cdot\frac1\rho^2 G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy - 
({\bf t}_\xx\cdot\nabla_\zz)\int_\Gamma \frac{\partial G(\zz-\yy)}{\partial s_\yy}\sigma(\yy)\ \dif s_\yy\\
&= -\frac1\rho^2 {\bf t}_\xx\cdot \int_\Gamma G(\zz-\yy){\bf t}_\yy \sigma(\yy)\ \dif s_\yy + 
({\bf t}_\xx \cdot \nabla_\zz)\int_\Gamma G(\zz-\yy)\frac{\partial \sigma(\yy)}{\partial s}\ \dif s_\yy.
\end{align}
%
Here ${\bf t}_\xx$ is the tangent vector at the point $\xx$ and $d/ds$ is the arclength derivative.
Letting $\zz\to\xx\in\Gamma$, we obtain
%
\begin{equation}
\label{eq:normal_deriv}
\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) = -\frac1\rho^2 {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx) + \frac{\partial}{\partial s}\SSS\left[\frac{\partial \sigma}{\partial s}\right](\xx).
\end{equation}
%

Substituting \eqref{eq:normal_deriv} to \eqref{eq:free_energy2} provides the total energy calculation
throughout this paper and this expression takes the advantage of not calculating the quadruple layer potential which has well-known obstacle in numerical implementation.

%Suppose we split the solution to the screened Laplace boundary value problem~\eqref{eq:SL}
%into smooth portion $w_i$ and singular portion $u_i$. That is, 
%\begin{equation}
%w_i({\bf x}) = u({\bf x}) - u_i = u({\bf x}) - \int_{\Gamma_i }\left(\pderiv{}{\nnu_\yy}
%    K_0 \left(\frac{|\xx - \yy|}{\rho}\right)\right) \sigma_i(\yy) \,
%    \dif s_\yy, \quad \xx \in \Gamma_i.
%\end{equation}
%Consider the case when $f(u)=u^2$ in~\eqref{eq:free_energy}, we can rewrite the free energy functional as 
%
%\begin{align}
%\label{eq:free_energy2}
%  E[u(\xx)] &= C\rho \sum_{i=1}^{N_b}\int_{\Gamma_i} u \frac{\partial u}{\partial \nnu}  \,\dif s
%= C\rho\sum_{i=1}^{N_b}\int_{\Gamma_i} u \left(\frac{\partial w_i}{\partial \nnu}+\frac{\partial u_i}{\partial \nnu} \right) \, \dif s, \quad \xx\in\Gamma_i.
%\end{align}
%%%%
%Follow the divergence theorem on the last term in the equation above and denote $g_i(x)$ as the boundary data on $\Gamma_i$, we obtain
%\begin{align}
%\label{eq:free_energy3}
%  E[u(\xx)] &= C\rho\sum_{i=1}^{N_b}\left(\int_{\Gamma_i} u \frac{\partial w_i}{\partial \nnu} \,\dif s  + \int_{\mathbb{R}^n\setminus \overline{U}_i} \left( \nabla u \cdot\nabla u_i + u\Delta u_i\right) \,\dif \xx\right)\\
%%
%&= C\rho\sum_{i=1}^{N_b}\left(\int_{\Gamma_i} u \frac{\partial w_i}{\partial \nnu} \,\dif s  + \int_{\mathbb{R}^n\setminus \overline{U}_i} u\Delta u_i\,\dif \xx
%+\int_{\Gamma_i}\left(u+\frac12\sigma\right)\frac{\partial u_i}{\partial \nnu} \,\dif s     
%- \int_{\mathbb{R}^n\setminus \overline{U}_i} u\Delta u_i\,\dif \xx \right)\\
%&= C\rho\sum_{i=1}^{N_b}\left(\int_{\Gamma_i} u \frac{\partial w_i}{\partial \nnu} + 
%\left(g_i+\frac12\sigma_i\right)\frac{\partial u_i}{\partial \nnu}   \,\dif s \right), \quad \xx \in \Gamma_i.
%\end{align}
%
%
%

%  &= C\rho\sum_{i=1}^{N_b}\left(\int_{\Gamma_i} g_i \frac{\partial w_i}{\partial \nnu} 
% +  \left(u_i+\frac12\sigma\right)  \frac{\partial g_i}{\partial \nnu} \,\dif s - \int_{\mathbb{R}^n\setminus \overline{U}_i}\Delta g_i  \left(u_i+\frac12\sigma\right)  - g_i \left(u_i+\frac12\sigma\right) \,\dif \xx\right) \\
% &= C\rho\sum_{i=1}^{N_b}\int_{\Gamma_i} g_i \frac{\partial w_i}{\partial \nnu}+ \left(u_i+\frac12\sigma\right)  \frac{\partial g_i}{\partial \nnu}\,\dif s, \quad \xx\in\Gamma_i.

%%
%$\sigma$ satisfies the second-kind integral equation
%\begin{equation}
%\label{eq:screenedSKIE}
%  g(\xx) = \frac{1}{2}\sigma(\xx) + 
%    \frac{1}{2\pi}\int_{\bd\Omega} \left(\pderiv{}{\nnu_\yy}
%    K_0 \left(\frac{|\xx - \yy|}{\rho}\right)\right) \sigma(\yy) \, 
%    \dif s_\yy, \quad \xx \in \bd\Omega,
%\end{equation}

\section{Numerical Results}

\subsection{Self-Assembled Janus Particle with Specific Boundary Conditions}

Distinct morphologies can be obtained from 
the HAP model by shifting the boundary condition $g(\xx)$.
Consider the linear case where $f(u) = u^2$.
This choice makes \eqref{eq:SL} a boundary value
problem for the screened-Laplace equation
and the solutions have a boundary layer structure that decays to $0$ in the
bulk with the decay length $\rho$.
We consider three kinds of boundary conditions
on particle $\Gamma_i$:
\begin{equation}
  \label{eq:bcs}
  \begin{tabular}{c|c|c}
     \text{(i)} & \text{(ii)} & \text{(iii)} \\
    \hline
    \rule{0pt}{4ex} 
      $g(\mathbf{x}) = \displaystyle \frac{1+\cos (\theta_i(\xx))}{\sqrt{3\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{2+\cos (\theta_i(\xx))}{\sqrt{9\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{\cos (\theta_i(\xx))}{\sqrt{\pi r}}$
\end{tabular}
\end{equation}
The number $\theta_i(\xx)$ is the angle formed by the position $\xx$,
the particle center $\aa_i$, and the particle director $\dd_i$.
(Figure~\ref{fig:flow_map}). The normalizations e.g., $(\pi r)^{-1/2}$,
provide a uniform surface energy $\int_{\Gamma_i} g^2 \,ds = 1$ for
circular particles of radius $r$. The cases (ii) and (iii) are obtained
by applying a vertical shift and scaling to case (i). Using this setup,
we simulate the dynamics of 60 particles with one set of random initial
positions and orientations as shown in Figure~\ref{fig:relax}(a)

\begin{figure}
  \begin{center}
  \includegraphics[width=0.2\textwidth]{LPA.jpg}
  \includegraphics[width=0.2\textwidth]{LPB.jpg}
  \includegraphics[width=0.2\textwidth]{LPC.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:bcs} Boundary conditions characterize the water
  structure at the particle interface: an amphiphilic particle (left), a
  hydrophobic particle with anisotropic intensity (middle), a water
  structure with positive/negative charge (right). The dashed curve is
  the boundary of the disk and the arrow is its director for each
  particle.}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{Fig2a.png}
  \includegraphics[width=0.4\textwidth]{Fig2b.png}\\
  \includegraphics[width=0.4\textwidth]{Fig2c.png}
  \includegraphics[width=0.4\textwidth]{Fig2d.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:relax} (a) Initial configuration for $N_b=60$
  structures. In order to expedite the self-assembly procedure, all
  particles are placed in a $8\times8$ box initially. (b)--(d)
  Equilibrium states for all boundary conditions~\eqref{eq:bcs}(i)--(iii), respectively. Colors represent the boundary values that increases from minimum (blue) to maximum (red).  Panel (b) reaches the equilibrium at $t=1200$; Panel (c) reaches the steady-state at $t=600$; Panel (d) reaches the final state at $t=180$.}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.5\textwidth]{Relax_Energy.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:relax_energy} Energy profile for all three boundary conditions.}
\end{figure}

Case (i) models amphiphilic particles.
Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and these are accounted for as follows.
The hydrophilic side takes the value $u =0$.
This mimics the apolar head of a lipid, for example, which does
not alter the structure of adjacent waters. The hydrophobic side
represents hydrocarbons and takes the value 
$u > 0$. The interaction between particles is attractive,
and particles will collectively orient their tails toward
one another. Fig~\ref{fig:relax}(b) demonstrates multiple bilayer structures such as pancake and vesicle shapes.

Multilamelar bilayers arise when both sides of the particle 
are hydrophobic. 
The boundary condition \eqref{eq:bcs}(ii) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (i).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads.%~\cite{Huetal19, deMeetal21}. 
Fig~\ref{fig:relax}(c) shows the multilamellar structure with 8 layers when $N_b=60$. 
The number of folds depends on several factors, for instances, number of
particles and initial configurations.

Finally, boundary condition \eqref{eq:bcs}(iii) models 
a particle whose head surface repels the tail surface as proposed in
\cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure
resembles a checkerboard pattern
where each particle coordinates
its head with the head of three other particles and its tail with the
tail of three other particles. % Fig~\ref{fig:self-assembly}(c).
Fig~\ref{fig:relax}(d) gives an example that this boundary condition results a checker board/stripes equilibrium state.

Fig~\ref{fig:relax_energy} shows the energy profile of boundary conditions (i)--(iii) using 
\eqref{eq:free_energy2} with \eqref{eq:normal_deriv} for normal derivative expression. As expected, for all simulation runs, the total energy of the systems decreases and reaches equilibrium states.

\subsection{Collective Janus Particles Suspended in a Shear Flow}
%%%

Consider placing the equilibrium state of boundary condition \eqref{eq:bcs} (ii) for 60 particles in a shear flow,
\begin{equation}
\uu_\infty(\xx) = \dot\gamma ({\bf e}_y\cdot\xx){\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the shear rate. Fig~\ref{fig:shear_1} (a)--(c) show the snapshots of simulation results when $\dot\gamma=0.05$.
We measure the ratio of major and minor axes $\lambda$ by adopting an ellipse fitting using the singular value decomposition and track this aspect ratio for this multi-lamellar structure when the flow is on. 
The result is shown in panel (d) and we observe that the multi-lamellar structure will be slightly stretched and the aspect ratio $\lambda=3.5$ increases over time.
%
Notice that for this simulation, the center of mass position of the
initial configuration is not at the origin and it is slightly to the
right of the origin. We also track the aspect ratio $\lambda$ for the
case that the center of mass position is shifted to the origin. Panel
(d) also shows that the dynamics of two sets of simulations are
invariant of center of mass position in a shear flow.

%
\begin{figure}
  \begin{center}
  \includegraphics[width=0.21\textwidth]{shear_0.png}
  \includegraphics[width=0.21\textwidth]{shear_1000.png}
  \includegraphics[width=0.21\textwidth]{shear_2000.png} 
  \includegraphics[width=0.21\textwidth]{shear_2000.png} \\
  \includegraphics[width=0.21\textwidth]{shear_checker_0.png}
  \includegraphics[width=0.21\textwidth]{shear_checker_250.png}
  \includegraphics[width=0.21\textwidth]{shear_checker_500.png}
  \includegraphics[width=0.21\textwidth]{shear_checker_750.png} \\
  \includegraphics[width=0.4\textwidth]{lambda_fig1.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:shear_1} Snapshots of simulation using boundary condition \eqref{eq:bcs}(ii) in a shear flow when $\dot\gamma=0.05$. }
\end{figure}
%

%Consider the energy profiles for two simulations where the centor of mass is on and off the origin.
%Fig~\ref{fig:shear_energy1} shows numerical results for both energy profiles and magnitude of energy differences is significantly large when there is a large deformation, for instance, the case when a multi-lamellar structure deformed to a pancake shape.
%
%\begin{figure}
%  \begin{center}
%  \includegraphics[width=0.5\textwidth]{shear_energy1.jpg}
%  \end{center}
%  \vspace{-20pt}  
%  \caption{\label{fig:shear_energy1} Energy profiles tracked during the simulations using boundary condition \eqref{eq:bcs}(ii) in a shear flow when $\dot\gamma=0.05$. }
%\end{figure}

%%%

The second set of simulation is putting the equilibrium state of the
boundary condition \eqref{eq:bcs} (iii), the checker board, in a shear
flow. After a set of empirical numerical investigations, we obtain a
critical shear rate $\dot\gamma=0.15$ where the checker board deforms
into stripes and small groups of JP. Figure~\ref{fig:shear_2}(a)--(d) show configurations when $t=\{0,50,100,150\}$. %The deformation starts with stripe separations
%
%\begin{figure}
%  \begin{center}
%  \includegraphics[width=0.4\textwidth]{shear_checker_0.png}
%  \includegraphics[width=0.4\textwidth]{shear_checker_250.png}\\
%  \includegraphics[width=0.4\textwidth]{shear_checker_500.png}
%    \includegraphics[width=0.4\textwidth]{shear_checker_750.png}
%  \end{center}
%  \vspace{-20pt}  
%  \caption{\label{fig:shear_2} Snapshots of simulation using boundary condition \eqref{eq:bcs}(iii) in a shear flow when $\dot\gamma=0.15$. }
%\end{figure}

\subsection{Collective Janus Particles Suspended in a Taylor-Green Flow}

Similar to shear flow simulations, we then place equilibrium states of all boundary conditions \eqref{eq:bcs} in a rotating Taylor-Green Flow with a length scale $l$,
\begin{equation}
\uu_\infty(\xx) = V_0 \left(-\cos(l{\bf e}_x\cdot\xx)\sin(l{\bf e}_y\cdot\xx){\bf e}_x+\sin(l{\bf e}_x\cdot\xx)\cos(l{\bf e}_y\cdot\xx){\bf e}_y\right).
\end{equation}
%
The magnitude of the vortex flow is controlled by the velocity scale
$V_0$.  Similar to the setup in the shear flow cases, we shift the
center of mass position to the origin then turn on the Taylor-Green flow
with $V_0=0.1$ and tune the sides of zero-velocity box to $\pi$, i.e.,
$l=0.5$, Figure~\ref{fig:TG_1}(a)--(c) show three frames of simulation
results and we track the aspect ratio $\lambda$ described in the
previous section which is provided in panel (d).  The stiffness of the
multi-lamellar structure is able to compete with the flow therefore the
structure does not deform to any states other than multi-lamellar shape.
We also found that the flow strength coefficient above $V_0=0.1$ will
result to clear group separations during the simulations.  Comparing
with the case when $V_0=0.2$, we provide a frame in
Figure~\ref{fig:TG_2}(a) and the energy comparison is shown in panel
(b). Numerical results validate the phenomenon that a higher
Taylor-Green vortex flow generates large structure separations with
higher total energy.

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{TG_0.png}
  \includegraphics[width=0.4\textwidth]{TG_2500.png}\\
  \includegraphics[width=0.4\textwidth]{TG_5000.png}
    \includegraphics[width=0.4\textwidth]{TG_dist.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:TG_1} Snapshots of simulation using boundary condition \eqref{eq:bcs}(b) in a Taylor-Green flow when $V_0=0.1$ and $l=0.5$.}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{TG0p2_2000.png}
  \includegraphics[width=0.4\textwidth]{TG_energy1.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:TG_2} (a) Snapshots of simulation using boundary condition \eqref{eq:bcs}(b) in a Taylor-Green flow when $V_0=0.2$, $l=0.5$, and $t = 400$. (b) Energy profiles for two cases $V_0=0.1$ and $V_0=0.2$. }
\end{figure}

\subsection{Steady-State Structure Confined in Narrow Channels}

Here are demonstrations of having JP vesicle and multi-lamellar
structure in a narrow channel under a shear flow.  Simulation results
and snapshots are shown in Figures~\ref{fig:confined_vesicle} and
Figure~\ref{fig:confined_lamellar} where the JP vesicle has $N_b=25$
particles and JP lamellar has $N_b=40$ particles, respectively. Both
starting configurations are pre-relaxed and the channel width in the
narrow part is about 5 length units in the vesicle case and 4 length
units in the lamellar case.

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{VesConf_0.png}
  \includegraphics[width=0.4\textwidth]{VesConf_2000.png}\\
  \includegraphics[width=0.4\textwidth]{VesConf_5000.png}
  \includegraphics[width=0.4\textwidth]{VesConf_10000.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:confined_vesicle} A JP vesicle confined in a narrow channel where $N_b=25$ and $\dot\gamma=0.01$.  }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{MultConf_100.png}
  \includegraphics[width=0.4\textwidth]{MultConf_600.png}\\
  \includegraphics[width=0.4\textwidth]{MultConf_1200.png}
  \includegraphics[width=0.4\textwidth]{MultConf_2400.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:confined_lamellar} A JP multi-lamellar structure confined in a narrow channel where $N_b=40$ and $\dot\gamma=0.05$.  }
\end{figure}

\section{Conclusions}

%%%

% If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}

% figures should be put into the text as floats.
% Use the graphics or graphicx packages (distributed with LaTeX2e)
% and the \includegraphics macro defined in those packages.
% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
% and Frank Mittelbach for instance.
%
% Here is an example of the general form of a figure:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Use the figure* environment if the figure should span across the
% entire page. There is no need to do explicit centering.

% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}

% Surround figure environment with turnpage environment for landscape
% figure
% \begin{turnpage}
% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}
% \end{turnpage}

% tables should appear as floats within the text
%
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
% The ruledtabular enviroment adds doubled rules to table and sets a
% reasonable default table settings.
% Use the table* environment to get a full-width table in two-column
% Add \usepackage{longtable} and the longtable (or longtable*}
% environment for nicely formatted long tables. Or use the the [H]
% placement option to break a long table (with less control than 
% in longtable).
% \begin{table}%[H] add [H] placement to break table across pages
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% Lines of table here ending with \\
% \end{tabular}
% \end{ruledtabular}
% \end{table}

% Surround table environment with turnpage environment for landscape
% table
% \begin{turnpage}
% \begin{table}
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% \end{tabular}
% \end{ruledtabular}
% \end{table}
% \end{turnpage}

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
%\appendix
%\section{}